---
title: "Metagenomics analysis workflow on MinION data"
author: "by PuntSeq: [Andre Holzer](https://orcid.org/0000-0003-2439-6364), [Lara Urban](https://orcid.org/0000-0002-5445-9314) and [Maximilian Stammnitz](https://orcid.org/0000-0002-1704-9199)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    theme: paper
    number_sections: false
    fig_width: 7
    fig_height: 6
    fig_caption: true
  pdf_document:
    toc: true
    number_sections: false
---


# Part 0: Initiation

This R markdown file allows to reproduce the downstream processing performed on minimap2 classifier outputs. It combines R, Python and Bash code and requires the following packages/dependencies to be installed on your computer before execution:

1.) R version 4.0.3 or higher. Read here (https://www.r-project.org) how to install.

2.) A version of conda. Please read here (https://conda.io/projects/conda/en/latest/user-guide/install/index.html?highlight=conda) how to install conda.

3.) The conda environment.yml file ("PuntSeq-analysis-env.yml") from our GitHub repository (https://github.com/d-j-k/puntseq). Create a local environment with the environment.yml file by executing the follwoing command from your shell.

```{bash eval=FALSE}
wget --no-check-certificate https://raw.githubusercontent.com/d-j-k/puntseq/master/conda_environment/PuntSeq-analysis-env.yaml
conda env create -f PuntSeq-analysis-env.yaml
```

## Installation & loading of required R packages
All packages which are required to execulte the anaylsis will be loaded and if not already done so first installed on your machine. The following lines of code will do both automatically for you. (In case there should be any errors while trying to insall any package, please perform installation of the required packages manually)
```{r}
list.of.packages = c("tidyverse","knitr","hash","RColorBrewer","scales", "ggpubr", "ggsci","FactoMineR", "factoextra","ggpmisc","stringr","reticulate","dendextend","pheatmap","grid","ggrepel","RColorBrewer","extrafont","readxl","ggbeeswarm","car","gameofthrones") 
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages) > 0) {install.packages(new.packages)}
lapply(list.of.packages, require, character.only=T)
```

## Global variables and functions
Next, global variables and functions are defined (Can be modify if nessesary)
```{r eval=FALSE}
# Date
date <- format(Sys.Date(), format="%Y-%m-%d")

# Set working directory 
setwd("./")

# Name of the output directory (will be created in your current working directory)
outfolder <- str_c(date,"_PuntSeq_Downstream-metagenomics-analysis/")

# Basecaller
bcaller <- "Guppy"  

# Classifier
classifyr <- "minimap" 

# Define which downsampling dataset will be used
dsampling <- 0  # 0: use orginal downsampling (by PuntSeq 2019-11-13), 1: generate new downsampling

# Color sheme for locations 1-9.1/9.2 + positiv & negativ control
lcolors <- c("1" = "#2157A4", "2" = "#3694D1", "3" = "#65C6E8", "4" = "#9DD7ED", "5" = "#D6DE4F", "6" = "#FDCB44", "7" = "#F1861E", "8" = "#E63A11", "9.1" = "#a50026", "9.2" = "#a50026", "Mock" = "#636363","P" = "#636363", "N" = "#f0f0f0")

# Color sheme for sampling timepoints in April, June and August 
tcolors <- c("April" = "#f0f0f0", "June" = "#bdbdbd", "August" = "#636363") # gerys
          #c("April" = "#66c2a5", "June" = "#fc8d62", "August" = "#8da0cb") # qualitative
          #c("April" = "#e0ecf4", "June" = "#9ebcda", "August" = "#8856a7") # sequential

# importing fonts
font_import()

# Scientfic scale function
scientific_1 <- function(x) {
  parse(text=gsub("e", " %*% 10^", scales::scientific_format()(x)))
}
```
and main directoriy for output data is created.
```{r eval=FALSE}
system(str_c("mkdir ",outfolder))
```

## Downloading classifier output data
Classifier outputs comprising raw count data for all samples obtained in April, June and August can directly be downloaded from our GitHub directory (https://github.com/d-j-k/puntseq). By executing the following lines of code this will automatically be performed saving the files to the "<OUTFOLDER>/Classifier_output" directory.
```{r eval=FALSE}
# Create sub-output directory
system(str_c("mkdir ",outfolder,"Classifier_output/"))

# download classifier output data on family level for april samples
infile_april_family <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/minimap2_family_april.txt")
destfile_april_family=str_c(outfolder,"Classifier_output/",bcaller,"_April_",classifyr,"_family.tsv")
if(!file.exists(destfile_april_family)){
  res <- tryCatch(download.file(infile_april_family, destfile_april_family, method="auto"), error=function(e) 1)
}

# download classifier output data on genus level for april samples
infile_april_genus <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/minimap2_genus_april.txt")
destfile_april_genus=str_c(outfolder,"Classifier_output/",bcaller,"_April_",classifyr,"_genus.tsv")
if(!file.exists(destfile_april_genus)){
  res <- tryCatch(download.file(infile_april_genus, destfile_april_genus, method="auto"), error=function(e) 1)
}

# download classifier output data on family level for june samples
infile_june_family <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/minimap2_family_june.txt")
destfile_june_family=str_c(outfolder,"Classifier_output/",bcaller,"_June_",classifyr,"_family.tsv")
if(!file.exists(destfile_june_family)){
  res <- tryCatch(download.file(infile_june_family, destfile_june_family, method="auto"), error=function(e) 1)
}

# download classifier output data on genus level for june samples
infile_june_genus <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/minimap2_genus_june.txt")
destfile_june_genus=str_c(outfolder,"Classifier_output/",bcaller,"_June_",classifyr,"_genus.tsv")
if(!file.exists(destfile_june_genus)){
  res <- tryCatch(download.file(infile_june_genus, destfile_june_genus, method="auto"), error=function(e) 1)
}

# download classifier output data on family level for august samples
infile_august_family <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/minimap2_family_august.txt")
destfile_august_family=str_c(outfolder,"Classifier_output/",bcaller,"_August_",classifyr,"_family.tsv")
if(!file.exists(destfile_august_family)){
  res <- tryCatch(download.file(infile_august_family, destfile_august_family, method="auto"), error=function(e) 1)
}

# download classifier output data on genus level for august samples
infile_august_genus <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/minimap2_genus_august.txt")
destfile_august_genus=str_c(outfolder,"Classifier_output/",bcaller,"_August_",classifyr,"_genus.tsv")
if(!file.exists(destfile_august_genus)){
  res <- tryCatch(download.file(infile_august_genus, destfile_august_genus, method="auto"), error=function(e) 1)
}

# download total read numbers for all timepoints and locations
infile_root <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/metadata/Number_classified_reads_minimap2.txt")
destfile_root=str_c(outfolder,"Classifier_output/",bcaller,"_",classifyr,"_root_april-june-august.tsv")
if(!file.exists(destfile_root)){
  res <- tryCatch(download.file(infile_root, destfile_root, method="auto"), error=function(e) 1)
}
```

## Downloading additional supplementary and metadata
Similarly, metadata and supplementary tables are downloaded from our GitHub directory (https://github.com/d-j-k/puntseq) to the "<OUTFOLDER>/Classifier_output" directory.
```{r eval=FALSE}
# Create sub-output directory
system(str_c("mkdir ",outfolder,"Metadata/"))

# download metadata
infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/metadata/Metadata_guppy_minimap2_table2.txt")
destfile_metadata=str_c(outfolder,"Metadata/Metadata_guppy_minimap2_table2.txt")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

# Create sub-output directory
system(str_c("mkdir ",outfolder,"Supplementary_data/"))

# download supplementary_data for benchmarking
infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/2020-11-03_tool_benchmark_resources.xlsx")
destfile_metadata=str_c(outfolder,"Supplementary_data/2020-11-03_tool_benchmark_resources.xlsx")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/2020-01-31_Minimap2_results_family_level_observed_vs_expected_full.tsv")
destfile_metadata=str_c(outfolder,"Supplementary_data/2020-01-31_Minimap2_results_family_level_observed_vs_expected_full.tsv")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/2020-01-31_Minimap2_results_family_level_observed_vs_expected_withoutEntero.tsv")
destfile_metadata=str_c(outfolder,"Supplementary_data/2020-01-31_Minimap2_results_family_level_observed_vs_expected_withoutEntero.tsv")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

# download supplementary_data for hydrochemical analysis
infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/Gaillardet_1999_Table1.txt")
destfile_metadata=str_c(outfolder,"Supplementary_data/Gaillardet_1999_Table1.txt")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

# download supplementary_data for clöassifier comparison
infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/Fig_S1_b_table.xlsx")
destfile_metadata=str_c(outfolder,"Supplementary_data/Fig_S1_b_table.xlsx")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/summary_poscon_april_family_majvote.tsv")
destfile_metadata=str_c(outfolder,"Supplementary_data/summary_poscon_april_family_majvote.tsv")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/summary_poscon_june_family_majvote.tsv")
destfile_metadata=str_c(outfolder,"Supplementary_data/summary_poscon_june_family_majvote.tsv")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/summary_poscon_august_family_majvote.tsv")
destfile_metadata=str_c(outfolder,"Supplementary_data/summary_poscon_august_family_majvote.tsv")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/summary_sample8_rep1_family_majority_vote.tsv")
destfile_metadata=str_c(outfolder,"Supplementary_data/summary_sample8_rep1_family_majority_vote.tsv")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/summary_sample8_rep2_family_majority_vote.tsv")
destfile_metadata=str_c(outfolder,"Supplementary_data/summary_sample8_rep2_family_majority_vote.tsv")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}

infile_metadata <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/supplementary_data/summary_sample8_rep3_family_majority_vote.tsv")
destfile_metadata=str_c(outfolder,"Supplementary_data/summary_sample8_rep3_family_majority_vote.tsv")
if(!file.exists(destfile_metadata)){
  res <- tryCatch(download.file(infile_metadata, destfile_metadata, method="auto"), error=function(e) 1)
}
```

## Initialise Python engine & conda environment
To execute the code junks written in phyton we recommend to run knit python engine together with our conda enviornment.
```{r eval=FALSE}
#devtools::install_github("rstudio/reticulate") # install the developer version in case there are any errors with activating the conda environment
library(reticulate)
use_condaenv("PuntSeq-analysis-env", required = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)
py_config()
```

# Part 1: Downsampling & Quality Control
## Continuous rarefactions of 16S classifications
### Load data
```{r eval=FALSE}
# a. Classifications with minimap2 vs. SILVA 16S (release 132)
classifications.genus <- classifications.family <- vector(mode = 'list', length = 3)
names(classifications.genus) <- names(classifications.family) <- c("April", "June", "August")
classifications.family$April <- read.table(destfile_april_family, sep = '\t', check.names = F, header = T, fill = T)
classifications.family$June <- read.table(destfile_june_family, sep = '\t', check.names = F, header = T, fill = T)
classifications.family$August <- read.table(destfile_august_family, sep = '\t', check.names = F, header = T, fill = T)
classifications.genus$April <- read.table(destfile_april_genus, sep = '\t', check.names = F, header = T, fill = T)
classifications.genus$June <- read.table(destfile_june_genus, sep = '\t', check.names = F, header = T, fill = T)
classifications.genus$August <- read.table(destfile_august_genus, sep = '\t', check.names = F, header = T, fill = T)

# b. Standardise column naming and order, remove "uncultured" and "unknown" taxa
classifications.family <- lapply(classifications.family, function(x){colnames(x)[1] <- 'Name'; ind <- match(c('Name', '1', '2', '3', '4', '5', '6', '7', '8', '9.1', '9.2', 'N', 'P'), colnames(x));
ind <- ind[!is.na(ind)]; x <- x[,ind]; colnames(x)[grep('P',colnames(x))] <- 'Mock'; x <- x[-grep('^Unknown Family$|^uncultured$',x[,'Name']),]; return(x)})

classifications.genus <- lapply(classifications.genus, function(x){colnames(x)[1] <- 'Name'; ind <- match(c('Name', '1', '2', '3', '4', '5', '6', '7', '8', '9.1', '9.2', 'N', 'P'), colnames(x));
ind <- ind[!is.na(ind)]; x <- x[,ind]; colnames(x)[grep('P',colnames(x))] <- 'Mock'; x <- x[-grep('^uncultured$',x[,'Name']),]; return(x)})

```

### Sampling functions
```{r eval=FALSE}
# a. Sample
sample_classifications <- function(x, n_rep, read_steps){
  
  if(nrow(x) > 0){
    
    # subsample multinomially for each
    sampling.x <-  do.call(cbind, lapply(read_steps, function(y) rmultinom(n = n_rep, size = y, prob = x[,2])))
    sampling.x.sets <- cbind(as.character(x[,1]),sampling.x)
    sampling.x.sets <- sampling.x.sets[!apply(sampling.x, 1, function(x){all(x == 0)}),]
    
    # calculate sats
    x <- cbind(x, sampling.x)
    summary.x <- t(rbind(mapply(function(ind) {get_richness_shannon(x = x[,c(1,ind)])}, ind = 3:ncol(x)),
                         mapply(function(ind) {get_simpson(x = x[,c(1,ind)])}, ind = 3:ncol(x))))
    
    # output
    summary.x <- matrix(unlist(summary.x), nrow = nrow(summary.x), ncol = 4)
    return(list(stats = summary.x, sets = sampling.x.sets))
    
  }else{
    
    # output
    return(matrix(NA, nrow = length(read_steps), ncol = 4))
    
  }
  
}

# b. Calculate sample richness, Shannon diversity index and evenness
get_richness_shannon <- function(x){
  
  # a. remove NAs/0s (if present)
  if(any(x[,2] == 0 | is.na(x[,2]))){
    x <- x[-which(x[,2] == 0 | is.na(x[,2])),]
  }
  
  # b. individuals in pool
  total_N <- sum(as.numeric(as.character(x[,2])))
  
  # c. type richness
  richness <- nrow(x)
  
  # d. calculate shannon index: iterate over each species/genus/order/phylum
  shannon_H <- -sum(apply(x, 1, function(y) {p_i <- as.numeric(y[2])/total_N; shannon_H <- p_i * log(p_i); return(shannon_H)}))
  
  # e. calculate maximum possible Shannon H
  shannon_H_max <- log(richness)
  
  # f. calculate Shannon E (evenness)
  shannon_E <- shannon_H/shannon_H_max
  
  # g. output
  out <- list("Richness" = richness,
              "Shannon H" = shannon_H,
              "Shannon E" = shannon_E)
  return(out)
}

# c. Calculate Simpson's D
get_simpson <- function(x){
  
  # a. remove NAs/0s (if present)
  if(any(x[,2] == 0 | is.na(x[,2]))){
    x <- x[-which(x[,2] == 0 | is.na(x[,2])),]
  }
  
  # b. individuals in pool
  total_N <- sum(as.numeric(as.character(x[,2])))
  
  # c. type richness
  richness <- nrow(x)
  
  # d. Simpson's D
  ## adding a pseudocount to account for species presence of 1 read (!)
  simpson_D <- 1 - sum(apply(x, 1, function(y) {p_i <- c(c(as.numeric(y[2])+1)*as.numeric(y[2]))/c(c(total_N+1)*c(total_N)); return(p_i)}))
  
  # e. output
  out <- list("Simpson D" = simpson_D)
  return(out)
}

# d. Generate Michaelis Menten fit do study downsampling effects
mmfit <- function(x, xlims, type){
  
  # take values, dose = counts, response = richness
  if (type == 'genus'){
    x.counts <- x[-c(nrow(x)-3,nrow(x)-2,nrow(x)-1,nrow(x)),1]
    x.richness <- x[-c(nrow(x)-3,nrow(x)-2,nrow(x)-1,nrow(x)),2]
  }else if (type == 'family'){
    x.counts <- x[-c(nrow(x)-4,nrow(x)-2,nrow(x)-1,nrow(x)),1]
    x.richness <- x[-c(nrow(x)-4,nrow(x)-2,nrow(x)-1,nrow(x)),6]  
  }
  
  if (length(x.counts) > 3 & length(unique(x.richness)) != 1){
    
    # MM model
    datas <- data.frame(x.counts, x.richness)
    datas <- datas[!is.na(datas[,2]),]
    
    MMcurve <- formula(x.richness ~ Vmax*x.counts/(Km+x.counts))
    bestfit <- nls(formula = MMcurve, 
                   data = datas, 
                   start = list(Vmax = max(datas[,2]),
                                Km = max(datas[,1])/2))
    SconcRange <- seq(f = 0, t = xlims[2], by = 10)
    theorLine <- predict(bestfit, list(x.counts = SconcRange))
    
    # Output
    out <- list(summary(bestfit)$parameters[,1], xy = cbind(SconcRange, theorLine))
    return(out)
  }
  
}
```
 
### Rarefaction curves for Genus and Family 
This section is only nessesary if a new subsampling should be performed.
```{r eval=FALSE}
if (dsampling == 1){
  
  # a. Parameters
  
  ## Number of reads
  read_steps <- seq(f = 100, t = 1800000, by = 100)
  
  ## Bootstraps per sampling depth
  n_rep <- 1
  
  ## Minimum number of supporting reads per genus/family/order/class/phylum for richness/shannon/simpson indices
  min_reads <- 5
  
  # c. Iterative sampling
  
  ## Prepare summary matrix
  sampling <- vector(mode = 'list', length = 12)
  names(sampling) <- c('1', '2', '3', '4', '5', '6', '7', '8', '9.1', '9.2', 'N', 'Mock')
  for (i in 1:length(sampling)){
    sampling[[i]] <- vector(mode = 'list', length = 3)
    names(sampling[[i]]) <- c("April", "June", "August")
    sampling[[i]]$April <- data.frame(matrix(NA, ncol = 9, nrow = length(read_steps)*n_rep))
    colnames(sampling[[i]]$April) <- c('Counts used',
                                       'Genus richness', 'Genus Shannon-H', 'Genus Shannon-E', 'Genus Simpson-D', 
                                       'Family richness', 'Family Shannon-H', 'Family Shannon-E', 'Family Simpson-D')
    sampling[[i]]$August <- sampling[[i]]$June <- sampling[[i]]$April
  }
  
  ## Iterate over each Cam site
  for (barcode in 1:length(sampling)){
    
    # Iterate over each time point (April, June, August) per Cam site
    for (time in 1:length(sampling[[barcode]])){
      
      ## Where we at, ma'es ?
      print(paste0('Processing sample ', names(sampling)[barcode], ' - ', names(classifications.family)[time]))
      
      ## Skip if data is yet missing
      barcode.name.tmp <- names(sampling)[barcode]
      out.ind <- grep(paste0('^',barcode.name.tmp,'$'), colnames(classifications.family[[time]]))
      if(length(out.ind) == 0){
        next
      }
      
      sampling[[barcode]][[time]][, 'Counts used'] <- rep(sort(rep(read_steps, n_rep)))
      
      ## fetch data
      tmp.data.family <- classifications.family[[time]][,c(1,out.ind)]
      tmp.data.genus <- classifications.genus[[time]][,c(1,out.ind)]
      
      ## sample multinomially on genus and family level
      tmp.data.genus <- tmp.data.genus[which(tmp.data.genus[,2] >= min_reads), ,drop = F]
      tmp.data.family <- tmp.data.family[which(tmp.data.family[,2] >= min_reads), ,drop = F]
      
      ## only take samplings which would lie below the actual number of classified reads (avoid UPsampling)
      sampling_thresh <- max(sum(tmp.data.genus[,2], na.rm = T), sum(tmp.data.family[,2], na.rm = T))
      sampling[[barcode]][[time]] <- sampling[[barcode]][[time]][!sampling[[barcode]][[time]][,1] > sampling_thresh,,drop = F]
      
      if(dim(tmp.data.genus)[1] > 0 & sum(tmp.data.genus[,2]) > read_steps[1]){
        read_steps.tmp <- seq(f = 100, t = max(sampling[[barcode]][[time]][,1]), by = 100)
        
        ## also generate richness & shannon indeces for each sample
        sampling[[barcode]][[time]][,c(2:5)] <- sample_classifications(x = tmp.data.genus, n_rep = n_rep, read_steps = read_steps.tmp)$stats
        sampling[[barcode]][[time]][,c(6:9)] <- sample_classifications(x = tmp.data.family, n_rep = n_rep, read_steps = read_steps.tmp)$stats
      }
  
      ## set samplings which would lie higher above taxon-level wise number of classified reads
      if(nrow(sampling[[barcode]][[time]]) > 0){
        sampling[[barcode]][[time]][sampling[[barcode]][[time]][,1] > sum(tmp.data.genus[,2], na.rm = T),2:5] <- NA
        sampling[[barcode]][[time]][sampling[[barcode]][[time]][,1] > sum(tmp.data.family[,2], na.rm = T),6:9] <- NA
      }
      
      ## at the end of the summary matrix, add original classification stats for genus, order and phylum
      sampling[[barcode]][[time]] <- rbind(sampling[[barcode]][[time]],
                                           rep(NA, ncol(sampling[[barcode]][[time]])),
                                           rep(NA, ncol(sampling[[barcode]][[time]])))
      colnames(sampling[[barcode]][[time]]) <- c('Counts used',
                                                 'Genus richness', 'Genus Shannon-H', 'Genus Shannon-E', 'Genus Simpson-D',
                                                 'Family richness', 'Family Shannon-H', 'Family Shannon-E', 'Family Simpson-D')
      sampling[[barcode]][[time]][nrow(sampling[[barcode]][[time]])-1,1] <- sum(tmp.data.genus[,2], na.rm = T)
      sampling[[barcode]][[time]][nrow(sampling[[barcode]][[time]])-1,2:4] <- as.numeric(get_richness_shannon(tmp.data.genus[,c(1,2)]))
      sampling[[barcode]][[time]][nrow(sampling[[barcode]][[time]])-1,5] <- as.numeric(get_simpson(tmp.data.genus[,c(1,2)]))
      sampling[[barcode]][[time]][nrow(sampling[[barcode]][[time]]),1] <- sum(tmp.data.family[,2], na.rm = T)
      sampling[[barcode]][[time]][nrow(sampling[[barcode]][[time]]),6:8] <- as.numeric(get_richness_shannon(tmp.data.family[,c(1,2)]))
      sampling[[barcode]][[time]][nrow(sampling[[barcode]][[time]]),9] <- as.numeric(get_simpson(tmp.data.family[,c(1,2)]))
    }
  }
}
```

### Summarise information: total data vs. 37,000 reads
```{r eval=FALSE}

# i. Create output directories
system(str_c("mkdir ",outfolder,"Downsampling/"))
system(str_c("mkdir ",outfolder,"Downsampling/genus"))
system(str_c("mkdir ",outfolder,"Downsampling/family"))

# ii. Download original downsampling dataset if required
if (dsampling == 0){
  # download original rarefaction sampling data
  infile_rarefaction <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/rarefied_data/2019-11-13_guppy_minimap2_sampling.Rdata")
  destfile_rarefaction=str_c(outfolder,"Downsampling/2019-11-13_guppy_minimap2_sampling.Rdata")
  if(!file.exists(infile_rarefaction)){
  res <- tryCatch(download.file(infile_rarefaction, destfile_rarefaction, method="auto"), error=function(e) 1)
  }
  load(destfile_rarefaction)

  # download all files
  for (level in c("family","genus")){
    for (location in c("1","2","3","4","5","6","7","8","9.1","9.2","Mock","N")){
      for (Time in c("April","June","August")){
        infile_tmp <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/rarefied_data/2019-11-13_guppy_minimap2_sampling_37k/",level,"/",location,"-",Time,"_",level,".tsv")
        destfile_tmp=str_c(outfolder,"Downsampling/",level,"/",location,"-",Time,"_",level,".tsv")
        if(!file.exists(infile_tmp)){
        res <- tryCatch(download.file(infile_tmp, destfile_tmp, method="auto"), error=function(e) 1)
        }
      }
    }
  }
}

# iii Save new downsampling dataset if required
if (dsampling == 1){
  save(sampling, file = str_c(outfolder,"Downsampling/",date,"_sampling_",bcaller,"_",classifyr,".Rdata"))
}


# a. Create summary tables (on genus/family taxon level)
summary.genus <- matrix(NA, 3*12, 4)
colnames(summary.genus) <- c('Sampling depth', 'Sampling richness', 
                             'Richness at depth 37000X', 'Richness at depth 37000X (% orig.)')
rownames(summary.genus) <- sort(c(paste(names(sampling), '- 1April'), 
                                  paste(names(sampling), '- 2June'),
                                  paste(names(sampling), '- 3August')))
summary.genus <- summary.genus[c(1:30,34:36,31:33),]
rownames(summary.genus) <- gsub('1A|3A', 'A',rownames(summary.genus))
rownames(summary.genus) <- gsub('2J', 'J',rownames(summary.genus))
summary.family <- summary.genus
count <- 0
for (i in 1:12){
  for (j in 1:3){
    count <- count + 1
    
    # see if downsampling has already been done
    if(all(is.na(sampling[[i]][[j]]) == T)){next}
    summary.genus[count,1:2] <- as.numeric(sampling[[i]][[j]][nrow(sampling[[i]][[j]])-1,1:2])
    summary.family[count,1:2] <- as.numeric(sampling[[i]][[j]][nrow(sampling[[i]][[j]]),c(1,6)])

    if(summary.genus[count,1] > 37000){
      out <- mmfit(sampling[[i]][[j]], xlims = c(0,80000), type = 'genus')
      summary.genus[count,3] <- round(out$xy[which(out$xy[,1] == 37000),2],1)
      summary.genus[count,4] <- round(100*summary.genus[count,3]/summary.genus[count,2],1)
    }
    if(summary.family[count,1] > 37000){
      out <- mmfit(sampling[[i]][[j]], xlims = c(0,80000), type = 'family')
      summary.family[count,3] <- round(out$xy[which(out$xy[,1] == 37000),2],1)
      summary.family[count,4] <- round(100*summary.family[count,3]/summary.family[count,2],1)
    }
    
  }
}

# b. Repeat downsampling to fixed depth of 37,000 (on genus/family taxon level)
sample_classifications_single <- function(x, sample, n_rep, read_steps, min_reads){
  
  # a. pre-process
  x <- x[x[,sample,drop = F] >= min_reads,,drop = F]
  
  # b. run sampling function
  if(nrow(x) > 0){
    out <- sample_classifications(x = x[,c(1,sample)],
                                  n_rep = n_rep,
                                  read_steps = read_steps)
    out.stats <- cbind(read_steps, out$stats)
    colnames(out.stats) <- c('Depth', "Richness", "Shannon H", "Shannon E", 'Simpson D')
    return(list(stats = out.stats, sets = out$sets)) 
  }else{
    return(NA)
  }
  
}
genus.37000 <- vector(mode = 'list', length = 12)
names(genus.37000) <- c('1', '2', '3', '4', '5', '6', '7', '8', '9.1', '9.2', 'N', 'Mock')
family.37000 <- genus.37000
for (i in 1:length(sampling)){
  print(i)
  genus.37000[[i]] <- vector(mode = 'list', length = 3)
  names(genus.37000[[i]]) <- c("April", "June", "August")
  family.37000[[i]] <- genus.37000[[i]]
  for (j in 1:3){
    
    # see if downsampling has already been done
    barcode.name.tmp <- names(sampling)[i]
    out.ind <- grep(paste0('^',barcode.name.tmp,'$'), colnames(classifications.family[[j]]))
    if(length(out.ind) == 0){
      next
    }    
    genus.37000[[i]][[j]] <- sample_classifications_single(x = classifications.genus[[j]], 
                                                           sample = out.ind, 
                                                           n_rep = 100, 
                                                           read_steps = 37000,
                                                           min_reads = 5)
    
    family.37000[[i]][[j]] <- sample_classifications_single(x = classifications.family[[j]],
                                                            sample = out.ind, 
                                                            n_rep = 100, 
                                                            read_steps = 37000, 
                                                            min_reads = 5)
  }
}

# c. Output downsamplings as .TSVs
for (i in 1:12){
  print(i)
  for (j in 1:3){
    
    if(all(is.na(sampling[[i]][[j]]) == T) | nrow(sampling[[i]][[j]]) < 3){next}
    write.table(genus.37000[[i]][[j]]$sets, paste0(outfolder,'Downsampling/genus/', names(genus.37000)[i],'-', names(genus.37000[[i]])[j],'_genus.tsv'),
                quote = F, col.names = T, row.names = F, sep = '\t')
    
    write.table(family.37000[[i]][[j]]$sets, paste0(outfolder,'Downsampling/family/', names(family.37000)[i],'-', names(family.37000[[i]])[j],'_family.tsv'),
                quote = F, col.names = T, row.names = F, sep = '\t')
  }
}
```

### Rarefaction Plots for 16S classifications  
```{r eval=FALSE}
# 2. Plotting functions
#######################

# a. Richness plots
richness.plots <- function(x, type, fixed.xmax, fixed.ymax, real.samples, title){
  
  # Define plot ranges
  if(fixed.xmax == ''){
    xlims <- c(0, max(x$April[,"Counts used"],
                      x$June[,"Counts used"],
                      x$August[,"Counts used"], na.rm = T))
  }else{
    xlims <- c(0, fixed.xmax)
  }
  
  if (type == "genus"){
    ind.richness <- 2
    ylabs <- 'Genera'
    
  }else if (type == "family"){
    ind.richness <- 6
    ylabs <- 'Families'
    
  }else if (type == "order"){
    ind.richness <- 10
    ylabs <- 'Orders'
    
  }else if (type == "class"){
    ind.richness <- 14
    ylabs <- 'Classes'
    
  }else if (type == "phylum"){
    ind.richness <- 18
    ylabs <- 'Phyla'
    
  }
  
  if(fixed.ymax == ''){
    ylims <- c(0, max(x$April[,ind.richness],
                      x$June[,ind.richness],
                      x$August[,ind.richness], na.rm = T)) 
  }else{
    ylims <- c(0, fixed.ymax)
  }
  
  ## Plot April
  plot(x = x$April[,"Counts used"],
       y = x$April[,ind.richness],
       ylab = ylabs, xlab = '', xaxt = 'n',
       pch = 16, type = 'p', col = 'goldenrod1', cex = 0.5,
       xlim = xlims, ylim = ylims,
       main = paste0(title, ": Richness"), cex.main = 3, cex.lab = 1.7, cex.axis = 1.3) 
  
  ## June
  points(x = x$June[,"Counts used"],
         y = x$June[,ind.richness],
         pch = 16,
         col = 'darkorange',
         type = 'p', cex = 0.5)
  
  ## August
  points(x = x$August[,"Counts used"],
         y = x$August[,ind.richness],
         pch = 16,
         col = 'red',
         type = 'p', cex = 0.5)
  
  axis(side = 1, at = seq(f = 0, t = xlims[2], length.out = 5), labels = seq(f = 0, t = xlims[2], length.out = 5))
  
  ## add Michaelis-Menten values
  if(type == 'genus'){
    
    if(x$April[c(nrow(x$April)-4),1] > 37000){
      mm.april <- mmfit(x = x$April, xlims = xlims, type = type)[[1]]
    }else{
      mm.april <- c()
    }
    if(x$June[c(nrow(x$June)-4),1] > 37000){
      mm.june <- mmfit(x = x$June, xlims = xlims, type = type)[[1]]
    }else{
      mm.june <- c()
    }
    if(x$August[c(nrow(x$August)-4),1] > 37000){
      mm.august <- mmfit(x = x$August, xlims = xlims, type = type)[[1]]
    }else{
      mm.august <- c()
    }
    
  }else if(type == 'family'){
    
    if(x$April[c(nrow(x$April)-3),1] > 37000){
      mm.april <- mmfit(x = x$April, xlims = xlims, type = type)[[1]]
    }else{
      mm.april <- c()
    }
    if(x$June[c(nrow(x$June)-3),1] > 37000){
      mm.june <- mmfit(x = x$June, xlims = xlims, type = type)[[1]]
    }else{
      mm.june <- c()
    }
    if(x$August[c(nrow(x$August)-3),1] > 37000){
      mm.august <- mmfit(x = x$August, xlims = xlims, type = type)[[1]]
    }else{
      mm.august <- c()
    }
    
  }else if(type == 'order'){
    
    if(x$April[c(nrow(x$April)-2),1] > 37000){
      mm.april <- mmfit(x = x$April, xlims = xlims, type = type)[[1]]
    }else{
      mm.april <- c()
    }
    if(x$June[c(nrow(x$June)-2),1] > 37000){
      mm.june <- mmfit(x = x$June, xlims = xlims, type = type)[[1]]
    }else{
      mm.june <- c()
    }
    if(x$August[c(nrow(x$August)-2),1] > 37000){
      mm.august <- mmfit(x = x$August, xlims = xlims, type = type)[[1]]
    }else{
      mm.august <- c()
    }
    
  }else if(type == 'class'){
    
    if(x$April[c(nrow(x$April)-1),1] > 37000){
      mm.april <- mmfit(x = x$April, xlims = xlims, type = type)[[1]]
    }else{
      mm.april <- c()
    }
    if(x$June[c(nrow(x$June)-1),1] > 37000){
      mm.june <- mmfit(x = x$June, xlims = xlims, type = type)[[1]]
    }else{
      mm.june <- c()
    }
    if(x$August[c(nrow(x$August)-1),1] > 37000){
      mm.august <- mmfit(x = x$August, xlims = xlims, type = type)[[1]]
    }else{
      mm.august <- c()
    }
    
  }else if(type == 'phylum'){
    
    if(x$April[nrow(x$April),1] > 37000){
      mm.april <- mmfit(x = x$April, xlims = xlims, type = type)[[1]]
    }else{
      mm.april <- c()
    }
    if(x$June[nrow(x$June),1] > 37000){
      mm.june <- mmfit(x = x$June, xlims = xlims, type = type)[[1]]
    }else{
      mm.june <- c()
    }
    if(x$August[nrow(x$August),1] > 37000){
      mm.august <- mmfit(x = x$August, xlims = xlims, type = type)[[1]]
    }else{
      mm.august <- c()
    }
    
  }
  
  ## replot actual scores
  if(real.samples != 'n'){
    
    if (type == "genus"){
      highest.april <- nrow(x$April)-4
      highest.june <- nrow(x$June)-4
      highest.august <- nrow(x$August)-4
    }else if (type == "family"){
      highest.april <- nrow(x$April)-3
      highest.june <- nrow(x$June)-3
      highest.august <- nrow(x$August)-3
    }else if (type == "order"){
      highest.april <- nrow(x$April)-2
      highest.june <- nrow(x$June)-2
      highest.august <- nrow(x$August)-2
    }else if (type == "class"){
      highest.april <- nrow(x$April)-1
      highest.june <- nrow(x$June)-1
      highest.august <- nrow(x$August)-1
    }else if (type == "phylum"){
      highest.april <- nrow(x$April)
      highest.june <- nrow(x$June)
      highest.august <- nrow(x$August)
    }
    
    points(x = x$April[highest.april,"Counts used"],
           y = x$April[highest.april,ind.richness],
           pch = 21,
           col = 'black',
           bg = 'goldenrod1',
           type = 'p', cex = 2)
    
    points(x = x$June[highest.june,"Counts used"],
           y = x$June[highest.june,ind.richness],
           pch = 21,
           col = 'black',
           bg = 'darkorange',
           type = 'p', cex = 2)
    
    points(x = x$August[highest.august,"Counts used"],
           y = x$August[highest.august,ind.richness],
           pch = 21,
           col = 'black',
           bg = 'red',
           type = 'p', cex = 2) 
    
  }
  
  legend.text <- c()
  legend.col <- c()
  if (length(mm.april) != 0){
    legend.text <- c(legend.text, paste0('April, Rmax: ', round(mm.april[1], 0)))
    legend.col <- c(legend.col, 'goldenrod1')
  }
  if (length(mm.june) != 0){
    legend.text <- c(legend.text, paste0('June, Rmax: ', round(mm.june[1], 0)))
    legend.col <- c(legend.col, 'darkorange')
  }
  if (length(mm.august) != 0){
    legend.text <- c(legend.text, paste0('August, Rmax: ', round(mm.august[1], 0)))
    legend.col <- c(legend.col, 'red')
  }
  if (length(legend.text) != 0){
    legend("topleft", legend = legend.text, lwd = 2, 
           pch = 16, col = legend.col, bty = 'n', cex = 1.2)  
  }
  
}
richness.plots.spec <- function(x, type, fixed.xmax, fixed.ymax, real.samples, col){
  
  # Define plot ranges
  if(fixed.xmax == ''){
    xlims <- c(0, max(x[,"Counts used"], na.rm = T))
  }else{
    xlims <- c(0, fixed.xmax)
  }
  
  if (type == "genus"){
    ind.richness <- 2
    ylabs <- 'Taxonomic Genera'
    
  }else if (type == "family"){
    ind.richness <- 6
    ylabs <- 'Taxonomic Families'
    
  }else if (type == "order"){
    ind.richness <- 10
    ylabs <- 'Taxonomic Orders'
    
  }else if (type == "class"){
    ind.richness <- 14
    ylabs <- 'Taxonomic Classes'
    
  }else if (type == "phylum"){
    ind.richness <- 18
    ylabs <- 'Taxonomic Phyla'
    
  }
  
  if(fixed.ymax == ''){
    ylims <- c(0, max(x[,ind.richness], na.rm = T)) 
  }else{
    ylims <- c(0, fixed.ymax)
  }
  
  # Plot
  plot(x = x[,"Counts used"],
       y = x[,ind.richness],
       ylab = ylabs, xlab = 'Sampling Depth', xaxt = 'n',
       pch = 16, type = 'p', col = col, cex = 0.5,
       xlim = xlims, ylim = ylims,
       main = "", cex.main = 2, 
       cex.lab = 1.5, cex.axis = 1.2)
  
  axis(side = 1, at = seq(f = 0, t = xlims[2], length.out = 6), labels = seq(f = 0, t = xlims[2], length.out = 6))
  
  ## replot actual scores
  if(real.samples != 'n'){
    
    if (type == "genus"){
      highest.april <- nrow(x$April)-4
      highest.june <- nrow(x$June)-4
      highest.august <- nrow(x$August)-4
    }else if (type == "family"){
      highest.april <- nrow(x$April)-3
      highest.june <- nrow(x$June)-3
      highest.august <- nrow(x$August)-3
    }else if (type == "order"){
      highest.april <- nrow(x$April)-2
      highest.june <- nrow(x$June)-2
      highest.august <- nrow(x$August)-2
    }else if (type == "class"){
      highest.april <- nrow(x$April)-1
      highest.june <- nrow(x$June)-1
      highest.august <- nrow(x$August)-1
    }else if (type == "phylum"){
      highest.april <- nrow(x$April)
      highest.june <- nrow(x$June)
      highest.august <- nrow(x$August)
    }
    
    points(x = x$April[highest.april,"Counts used"],
           y = x$April[highest.april,ind.richness],
           pch = 21,
           col = 'black',
           bg = 'goldenrod1',
           type = 'p', cex = 2)
    
    points(x = x$June[highest.june,"Counts used"],
           y = x$June[highest.june,ind.richness],
           pch = 21,
           col = 'black',
           bg = 'darkorange',
           type = 'p', cex = 2)
    
    points(x = x$August[highest.august,"Counts used"],
           y = x$August[highest.august,ind.richness],
           pch = 21,
           col = 'black',
           bg = 'red',
           type = 'p', cex = 2) 
    
  }
  
}

# b. Shannon evenness plots
shannon.plots <- function(x, type, fixed.xmax, real.samples, title){
  
  # Define plot ranges
  if(fixed.xmax == ''){
    xlims <- c(0, max(x$April[,"Counts used"],
                      x$June[,"Counts used"],
                      x$August[,"Counts used"], na.rm = T))
  }else{
    xlims <- c(0, fixed.xmax)
  }
  
  if (type == "genus"){
    ind.shannon <- 4
    ylabs <- 'Genera'
    
  }else if (type == "family"){
    ind.shannon <- 8
    ylabs <- 'Families'
    
  }else if (type == "order"){
    ind.shannon <- 12
    ylabs <- 'Orders'
    
  }else if (type == "class"){
    ind.shannon <- 16
    ylabs <- 'Classes'
    
  }else if (type == "phylum"){
    ind.shannon <- 20
    ylabs <- 'Phyla'
    
  }
  ylims <- c(0, 1)
  
  # Plot
  plot(x = x$April[,"Counts used"],
       y = x$April[,ind.shannon],
       ylab = '', xlab = '', xaxt = 'n',
       pch = 16, type = 'p', col = 'goldenrod1', cex = 0.5,
       xlim = xlims, ylim = ylims,
       main = paste0(title, ": Shannon Evenness"), cex.main = 3, 
       cex.lab = 1.7, cex.axis = 1.3)
  
  ## June
  points(x = x$June[,"Counts used"],
         y = x$June[,ind.shannon],
         pch = 16,
         col = 'darkorange',
         type = 'p', cex = 0.5)
  
  ## August
  points(x = x$August[,"Counts used"],
         y = x$August[,ind.shannon],
         pch = 16,
         col = 'red',
         type = 'p', cex = 0.5)
  
  axis(side = 1, at = seq(f = 0, t = xlims[2], length.out = 5), labels = seq(f = 0, t = xlims[2], length.out = 5))
  
  ## replot actual scores
  if(real.samples != 'n'){
    
    if (type == "genus"){
      highest.april <- nrow(x$April)-4
      highest.june <- nrow(x$June)-4
      highest.august <- nrow(x$August)-4
    }else if (type == "family"){
      highest.april <- nrow(x$April)-3
      highest.june <- nrow(x$June)-3
      highest.august <- nrow(x$August)-3
    }else if (type == "order"){
      highest.april <- nrow(x$April)-2
      highest.june <- nrow(x$June)-2
      highest.august <- nrow(x$August)-2
    }else if (type == "class"){
      highest.april <- nrow(x$April)-1
      highest.june <- nrow(x$June)-1
      highest.august <- nrow(x$August)-1
    }else if (type == "phylum"){
      highest.april <- nrow(x$April)
      highest.june <- nrow(x$June)
      highest.august <- nrow(x$August)
    }
    
    points(x = x$April[highest.april,"Counts used"],
           y = x$April[highest.april,ind.shannon],
           pch = 21,
           col = 'black',
           bg = 'goldenrod1',
           type = 'p', cex = 2)
    
    points(x = x$June[highest.june,"Counts used"],
           y = x$June[highest.june,ind.shannon],
           pch = 21,
           col = 'black',
           bg = 'darkorange',
           type = 'p', cex = 2)
    
    points(x = x$August[highest.august,"Counts used"],
           y = x$August[highest.august,ind.shannon],
           pch = 21,
           col = 'black',
           bg = 'red',
           type = 'p', cex = 2) 
    
  }
  
}

# c. Simpson's D evenness plots
simpson.plots <- function(x, type, fixed.xmax, real.samples, title){
  
  # Define plot ranges
  if(fixed.xmax == ''){
    xlims <- c(0, max(x$April[,"Counts used"],
                      x$June[,"Counts used"],
                      x$August[,"Counts used"], na.rm = T))
  }else{
    xlims <- c(0, fixed.xmax)
  }
  
  if (type == "genus"){
    ind.simpson <- 5
  }else if (type == "family"){
    ind.simpson <- 9
  }else if (type == "order"){
    ind.simpson <- 13
  }else if (type == "class"){
    ind.simpson <- 17
  }else if (type == "phylum"){
    ind.simpson <- 21
  }
  ylims <- c(0, 1)
  
  # Plot
  plot(x = x$April[,"Counts used"],
       y = x$April[,ind.simpson],
       ylab = '', xlab = '', xaxt = 'n',
       pch = 16, type = 'p', col = 'goldenrod1', cex = 0.5,
       xlim = xlims, ylim = ylims,
       main = paste0(title, ": Simpson's D"), cex.main = 3, 
       cex.lab = 1.7, cex.axis = 1.3)
  
  ## June
  points(x = x$June[,"Counts used"],
         y = x$June[,ind.simpson],
         pch = 16,
         col = 'darkorange',
         type = 'p', cex = 0.5)
  
  ## August
  points(x = x$August[,"Counts used"],
         y = x$August[,ind.simpson],
         pch = 16,
         col = 'red',
         type = 'p', cex = 0.5)
  
  axis(side = 1, at = seq(f = 0, t = xlims[2], length.out = 5), labels = seq(f = 0, t = xlims[2], length.out = 5))
  
  ## replot actual scores
  if(real.samples != 'n'){
    
    if (type == "genus"){
      highest.april <- nrow(x$April)-4
      highest.june <- nrow(x$June)-4
      highest.august <- nrow(x$August)-4
    }else if (type == "family"){
      highest.april <- nrow(x$April)-3
      highest.june <- nrow(x$June)-3
      highest.august <- nrow(x$August)-3
    }else if (type == "order"){
      highest.april <- nrow(x$April)-2
      highest.june <- nrow(x$June)-2
      highest.august <- nrow(x$August)-2
    }else if (type == "class"){
      highest.april <- nrow(x$April)-1
      highest.june <- nrow(x$June)-1
      highest.august <- nrow(x$August)-1
    }else if (type == "phylum"){
      highest.april <- nrow(x$April)
      highest.june <- nrow(x$June)
      highest.august <- nrow(x$August)
    }
    
    points(x = x$April[highest.april,"Counts used"],
           y = x$April[highest.april,ind.simpson],
           pch = 21,
           col = 'black',
           bg = 'goldenrod1',
           type = 'p', cex = 2)
    
    points(x = x$June[highest.june,"Counts used"],
           y = x$June[highest.june,ind.simpson],
           pch = 21,
           col = 'black',
           bg = 'darkorange',
           type = 'p', cex = 2)
    
    points(x = x$August[highest.august,"Counts used"],
           y = x$August[highest.august,ind.simpson],
           pch = 21,
           col = 'black',
           bg = 'red',
           type = 'p', cex = 2) 
  }
}

# 3. Supplementary Figure: rarefaction effects on the Family level
##################################################################

pdf(str_c(outfolder,'Downsampling/',date,'_Supplementary_Figure_rarefactions.pdf'),
    width = 18, height = 7)
par(mfrow = c(1,2))
par(mar = c(7, 6, 5, 2))

## part A
richness.plots.spec(x = sampling[[1]]$April, type = 'family', 
                    fixed.xmax = 100000, fixed.ymax = 160, real.samples = 'n', col = '#2157A4')
abline(v = 37000, lty = 2, lwd = 0.5)

## part B
sampling.cols <- c('#2157A4', '#3694D1', '#65C6E8', '#9DD7ED', '#D6DE4F', 
                   '#FDCB44', '#F1861E', '#E63A11', '#a50026', '#a50026')
names(sampling.cols) <- c(1, 2, 3, 4, 5, 6, 7, 8, 9.1, 9.2)

## summarise richnesses at 37000X [you can add/remove samples here!]
tmp.rich <- cbind(family.37000$`1`$April$stats[,2],
                  family.37000$`1`$June$stats[,2],
                  family.37000$`1`$August$stats[,2],
                  family.37000$`2`$April$stats[,2],
                  family.37000$`2`$August$stats[,2], #
                  family.37000$`3`$April$stats[,2],
                  family.37000$`4`$April$stats[,2],
                  family.37000$`4`$August$stats[,2], #
                  family.37000$`5`$April$stats[,2],
                  family.37000$`5`$August$stats[,2],
                  family.37000$`6`$June$stats[,2],
                  family.37000$`6`$August$stats[,2],
                  family.37000$`7`$April$stats[,2],
                  family.37000$`8`$April$stats[,2],
                  family.37000$`9.1`$August$stats[,2],
                  family.37000$`9.2`$August$stats[,2])
colnames(tmp.rich) <- c('1 - April', 
                        '1 - June', 
                        '1 - August', 
                        '2 - April', 
                        '2 - August', 
                        '3 - April',
                        '4 - April', 
                        '4 - August', 
                        '5 - April', 
                        '5 - August', 
                        '6 - June', 
                        '6 - August',
                        '7 - April', 
                        '8 - April', 
                        '9.1 - August', 
                        '9.2 - August')

## fetch original richnesses
#genus.richness.orig <- summary.genus[match(colnames(tmp.rich), rownames(summary.genus)),2]
family.richness.orig <- summary.family[match(colnames(tmp.rich), rownames(summary.family)),2]

## obtain ratios and percentage values
for (i in 1:ncol(tmp.rich)){
  tmp.rich[,i] <- 100*c(tmp.rich[,i]/family.richness.orig[i])
}

## match colors in plot
col.match <- str_split_fixed(colnames(tmp.rich), ' - ', 2)[,1]
cols <- as.character(sampling.cols[match(col.match,names(sampling.cols))])

boxplot(tmp.rich,
        main = "", 
        cex.main = 2, ylab = 'Original richness at 37,000 X [%]', ylim = c(0, 100), 
        cex.lab = 1.5, cex.axis = 1, 
        pch = 16, cex = 0.3, notch = T, las = 2, yaxt = 'n',
        col = cols, 
        border = cols)
axis(2, at = seq(f = 0, t = 100, by = 20), las = 3, cex.axis = 1.2)
dev.off()


# 4. Three additional Figures (feel free to remove them)
########################################################

cols.all <- c(rep('#2157A4', 3), rep('#3694D1', 3), rep('#65C6E8', 3),
              rep('#9DD7ED', 3), rep('#F7EC73', 3), rep('#FDCB44', 3),
              rep('#F1861E', 3), rep('#E63A11', 3), rep('#D61015', 3), 
              rep('#D61015', 3), rep('grey', 3), rep('black', 3))

# a. Downsampling effects on richness and evenness for all samples
# pdf(outfolder,'Downsampling/',date,'_final_sampling_',classifyr,'_downsampling_richness_vs_evenness.pdf', 
#     width = 18, height = 7)
# par(mfcol = c(1,2))
# #for(type in c('genus', 'family', 'order', 'class', 'phylum')){
# for(type in c('family')){
# 
#   if(type == 'genus'){
#     fixed.ymax = 1000
#   }else if(type == 'family'){
#     fixed.ymax = 300
#   }else if(type == 'order'){
#     fixed.ymax = 250
#   }else if(type == 'class'){
#     fixed.ymax = 100
#   }else if(type == 'phylum'){
#     fixed.ymax = 50
#   }
#   
#   for (i in 1:12){
#     par(mar = c(4, 6, 4, 3))
#     richness.plots(x = sampling[[i]], type = type, fixed.xmax = 80000, fixed.ymax = fixed.ymax, real.samples = 'n',
#                    title = names(sampling)[i])
#     abline(v = 37000, lty = 2, lwd = 0.5)
#     shannon.plots(x = sampling[[i]], type = type, fixed.xmax = 80000, real.samples = 'n',
#                   title = names(sampling)[i])
#     abline(v = 37000, lty = 2, lwd = 0.5)
#   }  
# }
# dev.off()

# b. Original sequencing depths vs. 37,000 reads cutoff
pdf(str_c(outfolder,'Downsampling/',date,'_final_sampling_',classifyr,'_depth_cutoff_37000X.pdf'), 
    width = 8, height = 6)
# barplot(summary.genus[,1], las = 2, col = cols.all, border = cols.all, ylim = c(0,1000000), 
#         main = 'Original read depth (Genus)', cex.main = 2, cex.names = 0.7)
# abline(h = 37000, lty = 2)
barplot(summary.family[,1], las = 2, col = cols.all, border = cols.all, ylim = c(0,2000000), 
        main = 'Original read depth (Family)', cex.main = 2, cex.names = 0.7, cex.axis = 0.7)
abline(h = 37000, lty = 2)
# barplot(summary.order[,1], las = 2, col = cols.all, border = cols.all, ylim = c(0,1000000), 
#         main = 'Original read depth (Order)', cex.main = 2, cex.names = 0.7)
# abline(h = 37000, lty = 2)
# barplot(summary.class[,1], las = 2, col = cols.all, border = cols.all, ylim = c(0,1000000), 
#         main = 'Original read depth (Class)', cex.main = 2, cex.names = 0.7)
# abline(h = 37000, lty = 2)
# barplot(summary.phylum[,1], las = 2, col = cols.all, border = cols.all, ylim = c(0,1000000),
#         main = 'Original read depth (Phylum)', cex.main = 2, cex.names = 0.7)
# abline(h = 37000, lty = 2)
dev.off()

# c. Correlation of original sequencing depth vs. richness at 37000X
pdf(str_c(outfolder,'Downsampling/',date,'_final_sampling_',classifyr,'_richness_at_37000X_vs_sampling_depth.pdf'), width = 8, height = 6)

## Genus
plot(x = summary.family[1:30,1],
     y = summary.family[1:30,3], 
     pch = 16, 
     xlab = 'Original sampling Depth',
     ylab = 'Family richness at 37,000X',
     xlim = c(0,1000000),
     ylim = c(0,300), col = cols.all[1:30], 
     main = 'Family richness vs. 16S read depth', cex.main = 2)
pearson.cor <- round(cor(summary.family[1:30,1], summary.family[1:30,3], use = 'complete.obs'),3)
lm.out <- summary(lm(summary.family[1:30,3] ~ summary.family[1:30,1]))
pval <- round(lm.out[[4]][2,4],3)
abline(coef = lm.out[[4]][,1], lty = 2)
legend('topright', c(paste('Pearson correlation: ', pearson.cor),
                     paste('LM p-value: ', pval)))

# 
# ## Order
# plot(x = summary.order[1:30,1],
#      y = summary.order[1:30,3], 
#      pch = 16, 
#      xlab = 'Original sampling Depth',
#      ylab = 'Order richness at 37,000X',
#      xlim = c(0,1000000),
#      ylim = c(0,250), col = cols.all[1:30], 
#      main = 'Order richness vs. 16S read depth', cex.main = 2)
# pearson.cor <- round(cor(summary.order[1:30,1], summary.order[1:30,3], use = 'complete.obs'),3)
# lm.out <- summary(lm(summary.order[1:30,3] ~ summary.order[1:30,1]))
# pval <- round(lm.out[[4]][2,4],3)
# abline(coef = lm.out[[4]][,1], lty = 2)
# legend('topright', c(paste('Pearson correlation: ', pearson.cor),
#                      paste('LM p-value: ', pval)))
# 
# ## Class
# plot(x = summary.class[1:30,1],
#      y = summary.class[1:30,3], 
#      pch = 16, 
#      xlab = 'Original sampling Depth',
#      ylab = 'Class richness at 37,000X',
#      xlim = c(0,1000000),
#      ylim = c(0,100), col = cols.all[1:30], 
#      main = 'Class richness vs. 16S read depth', cex.main = 2)
# pearson.cor <- round(cor(summary.class[1:30,1], summary.class[1:30,3], use = 'complete.obs'),3)
# lm.out <- summary(lm(summary.class[1:30,3] ~ summary.class[1:30,1]))
# pval <- round(lm.out[[4]][2,4],3)
# abline(coef = lm.out[[4]][,1], lty = 2)
# legend('topright', c(paste('Pearson correlation: ', pearson.cor),
#                      paste('LM p-value: ', pval)))
# 
# ## Phylum
# plot(x = summary.phylum[1:30,1],
#      y = summary.phylum[1:30,3], 
#      pch = 16, 
#      xlab = 'Original sampling Depth',
#      ylab = 'Phylum richness at 37,000X',
#      xlim = c(0,1000000),
#      ylim = c(0,50), col = cols.all[1:30], 
#      main = 'Phylum richness vs. 16S read depth', cex.main = 2)
# pearson.cor <- round(cor(summary.phylum[1:30,1], summary.phylum[1:30,3], use = 'complete.obs'),3)
# lm.out <- summary(lm(summary.phylum[1:30,3] ~ summary.phylum[1:30,1]))
# pval <- round(lm.out[[4]][2,4],3)
# abline(coef = lm.out[[4]][,1], lty = 2)
# legend('topright', c(paste('Pearson correlation: ', pearson.cor),
#                      paste('LM p-value: ', pval)))
dev.off()
```

####  This section is only nessesary if a new subsampling should be performed
To replicate the results from the manuscript, use puntseq/minimap2_classifications/rarefied_data/final_family_1.tsv here: https://github.com/d-j-k/puntseq; final_family_2.tsv, final_family_3.tsv and final_family_4.tsv are the three additional downsampled datasets used in the manuscript; genus-level downsampled data is available in the same folder
```{python, eval=FALSE}
import datetime 
import pandas as pd
import io
import os
import requests
import numpy as np

DATE = datetime.datetime.now().strftime('%Y-%m-%d')
outfolder = DATE + '_PuntSeq_Downstream-metagenomics-analysis/'
bcaller = 'Guppy'
classifyr = 'minimap'

# build and save exemplary rarefaction dataframes
# example on the genus level
data = pd.read_csv(outfolder+'Downsampling/genus/%s' % os.listdir(outfolder+'Downsampling/genus/')[0], sep='\t', index_col=0, header=0) 
data = data[~data.index.duplicated(keep='first')]
data = data[data.taxRank == 'G']
del data['taxRank']
datacolumns = [os.listdir(outfolder+'Downsampling/genus/')[0]]
for filename in os.listdir(outfolder+'Downsampling/genus/')[1:]:
  if filename.endswith(".tsv"):
    data0 = pd.read_csv(outfolder+'Downsampling/genus/%s' %filename, sep='\t', index_col=0, header=0)
    data0 = data0[~data0.index.duplicated(keep='first')]
    data0 = data0[data0.taxRank == 'G']
    del data0['taxRank']
    data = pd.concat([data,data0], axis=1)
    datacolumns.append(filename)

# choose samples with more than 37k reads in the full dataset to keep for downstream analyses; in genus case:   -->
datacolumns_filt = ['1-April_genus.tsv', 
  '1-August_genus.tsv', 
  '1-June_genus.tsv', 
  '2-April_genus.tsv', 
  '2-August_genus.tsv', 
  '3-April_genus.tsv', 
  '4-April_genus.tsv',
  '4-August_genus.tsv', 
  '5-April_genus.tsv',
  '5-August_genus.tsv',
  '6-August_genus.tsv', 
  '6-June_genus.tsv', 
  '7-April_genus.tsv', 
  '8-April_genus.tsv', 
  '9.1-August_genus.tsv', 
  '9.2-August_genus.tsv', 
  'Mock-April_genus.tsv',
  'Mock-August_genus.tsv',
  'Mock-June_genus.tsv', 
  'N-August_genus.tsv']

posfilt = [np.where(np.array(datacolumns) == x)[0][0] for x in datacolumns_filt]

# create four exemplar rarefaction dataframes for later comparison; we will mostly work with "final_1.tsv" available on github (see above)
for nr in ['1','2','3','4']:
  dataa = data.copy()
  dataa = dataa[nr]
  dataa = dataa.dropna(how='all')
  dataa = dataa.fillna(0)
  dataa = dataa.iloc[:,posfilt]
  dataa.columns = datacolumns_filt
  dataa.to_csv(outfolder+'Downsampling/genus/final_%s.tsv' %nr, sep='\t')
```

### Mantel test
```{python, eval=FALSE}

import skbio
import pandas as pd
from sklearn.preprocessing import StandardScaler

# load one downsampled file, here on genus level
data = pd.read_csv('Downsampling/genus/final_1.tsv', sep='\t', index_col=0, header=0)

# process downsampled data
data = data.transpose()
data['month'] = [x.split('-')[1].split('_')[0] for x in data.index]
data['targetnice'] = [x.split('-')[0] for x in data.index]
data['targetnice'][data['targetnice'] == 'Mock'] = 'P'
data.index = pd.MultiIndex.from_arrays(data[['month', 'targetnice']].values.T, names=['month', 'location'])
del data['month']; del data['targetnice']

data = data.drop('N', level='location')
data = data.drop('P', level='location')

# load full genus (or family) minimap2 dataset available on github
url = "https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/All_genus_minimap2.tsv"
s = requests.get(url).content
fullall = pd.read_csv(io.StringIO(s.decode('utf-8'))', sep='\t', index_col=[0,1], header=0)

# process full data
fullall = fullall.iloc[:, np.where(fullall.columns.isin(data.columns))[0]]
fullall = fullall.loc[data.index.values]
data = data[fullall.columns]

# standadrise downscaled and full data
data = pd.DataFrame(StandardScaler().fit_transform(data), index = data.index, columns = data.columns)
fullall = pd.DataFrame(StandardScaler().fit_transform(fullall), index = fullall.index, columns = fullall.columns)

# calculate distance
datae = scipy.spatial.distance.cdist(data, data, metric='euclidean')
fullalle = scipy.spatial.distance.cdist(fullall, fullall, metric='euclidean')

# apply Mantel test to compare the distances of the full and downsampled data
skbio.stats.distance.mantel(datae, fullalle, method='pearson', permutations=99999, alternative='two-sided', strict=True, lookup=None)

```

##  Download/load downsampled classifier output data
```{r eval=FALSE}
# loop over all downsampling sets of both family and genus level
for (set in 1:4){
  for (level in c("family","genus")){
    
    # download and original downsampling files if nessesary
    if (dsampling == 0){
      # download downsampled datasets
      infile_downsampling<-str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/minimap2_classifications/rarefied_data/final_",level,"_",set,".tsv") 
      destfile_downsampling=str_c(outfolder,"Downsampling/",level,"/Downsampled_",level,"_set",set,"_",classifyr,".tsv")
      if(!file.exists(destfile_downsampling)){
        res <- tryCatch(download.file(infile_downsampling, destfile_downsampling, method="auto"), error=function(e) 1)
      }
    # else refer to new downsampled data files 
    } else {
      destfile_downsampling=str_c(outfolder,"Downsampling/",level,"/final_",level,"_",set,".tsv")
    }
    
    # define variable name
    var_name <- paste("classified_reads_downsampled_",level,"_set", set, sep = "")
    # assign data to variable
    assign(var_name, read_tsv(destfile_downsampling))
    
  }
}

# load downsampled datasets into lists
family.list <- list(classified_reads_downsampled_family_set1, classified_reads_downsampled_family_set2, classified_reads_downsampled_family_set3, classified_reads_downsampled_family_set4)
genus.list <- list(classified_reads_downsampled_genus_set1, classified_reads_downsampled_genus_set2, classified_reads_downsampled_genus_set3, classified_reads_downsampled_genus_set4)

# combine lists into a single multidimensional list
classified_reads_downsampled <- list(family.list, genus.list)
```

## Rawdata overview plots
### Pie charts of total reads per barcode for each MinION run
```{r}
# Create output directory
system(str_c("mkdir ",outfolder,"Overview-plots/"))

# Load data into dataframe and prepare for plotting
reads.data.raw <- read_tsv(str_c(outfolder,"Metadata/Metadata_guppy_minimap2_table2.txt"))
reads.data <- select(reads.data.raw, Barcode, Date, Reads, Reads_percent) %>%
  mutate(Date = str_replace(Date, pattern = "15.04.2018", "April")) %>%
  mutate(Date = str_replace(Date, pattern = "17.06.2018", "June")) %>%
  mutate(Date = str_replace(Date, pattern = "19.08.2018", "August")) %>%
  rename(barcode = Barcode) %>%
  rename(time = Date) %>%
  rename(n = Reads) %>%
  rename(perc = Reads_percent)

# Add label position
reads.data <- reads.data %>%
  mutate(lab.ypos = cumsum(perc) - 0.5*perc)

# Pie chart version 1
ggplot(reads.data, aes(x = "", y = perc, fill = barcode)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
  facet_grid(cols = vars(time))+ 
  geom_text(aes(label = paste0(round(perc), "%")), position = position_stack(vjust = 0.5), color = "white")+
  scale_fill_manual(values = lcolors) +
  theme_void()+
  theme(legend.position="right")

# save plot
ggsave(str_c(outfolder,"Overview-plots/",date,"_total_reads_per_barcod_",classifyr,"_1.png"), width = 20, height = 10, units = "cm")
ggsave(str_c(outfolder,"Overview-plots/",date,"_total_reads_per_barcod_",classifyr,"_1.pdf"), width = 20, height = 10, units = "cm") 

# Pie chart version 2
ggplot(reads.data, aes(x = 2, y = perc, fill = barcode)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0)+
  facet_grid(cols = vars(time))+ 
  geom_text(aes(label = paste0(round(perc), "%")), position = position_stack(vjust = 0.5), color = "white")+
  scale_fill_manual(values = lcolors) +
  theme_void()+
  xlim(0.5, 2.5)+
  theme(legend.position="right")

# save plot
ggsave(str_c(outfolder,"Overview-plots/",date,"_total_reads_per_barcod_",classifyr,"_2.png"), width = 20, height = 10, units = "cm")
ggsave(str_c(outfolder,"Overview-plots/",date,"_total_reads_per_barcod_",classifyr,"_2.pdf"), width = 20, height = 10, units = "cm") 
```

### Correlation analysis for DNA concentration and read numbers
```{r}
# Load data  from metadata file into dataframe and prepare for plotting
DNA.data.raw <- read_tsv(str_c(outfolder,"Metadata/Metadata_guppy_minimap2_table2.txt"))
DNA.data <- DNA.data.raw %>%
  mutate(Date = str_replace(Date, pattern = "15.04.2018", "April")) %>%
  mutate(Date = str_replace(Date, pattern = "17.06.2018", "June")) %>%
  mutate(Date = str_replace(Date, pattern = "19.08.2018", "August")) %>%
  transform(DNA_Concentration = as.numeric(DNA_Concentration)) %>%
  transform(DNA_Total = as.numeric(DNA_Total))  %>%
  rename(Time = Date)

# Correlation between DNA yield and PCR yield
 p1 <- ggscatter(DNA.data, x = "DNA_Concentration", y = "X16S_Concentration", xticks.by = 10)+
   geom_point(aes(color=Time))+
   scale_color_grey()+
   labs(title = "", caption = waiver(), x="DNA concentration (ng/µl)", y="16S concentration (ng/µl)")+
   geom_smooth(se = TRUE, method = "lm", formula = y ~ x, color = "black")+
   stat_cor(label.x = 2,label.y = 45, size = 4, method = "pearson", col="black") +           # Add correlation coefficient
   theme(legend.position="right")

 # Correlation between PCR yield and sequencing output
p2 <- ggscatter(DNA.data, x = "X16S_Concentration", y = "Reads", yscale = "log10")+
  geom_point(aes(color=Time))+
  labs(title = "", caption = waiver(), x="16S concentration (ng/µl)", y="Reads (1)")+
  scale_color_grey()+
  scale_y_continuous(trans='log10', breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x)))+
  annotation_logticks(sides="l")+
  geom_smooth(se = TRUE, method = "gam", formula = y ~ s(log(x)), color = "black")+
  stat_poly_eq(aes(color = Time, label = ..eq.label..), formula = y ~ s(log(x)), label.y = 0.9, label.x = 0.05, size = 2.5,parse = TRUE) +
  geom_vline(aes(xintercept=4), color="red",linetype="dashed")+
  theme(legend.position="top")

# combine both plots 
ggarrange(
  p1, p2, labels = c("A", "B"), ncol = 2, nrow = 1,
  common.legend = TRUE, legend = "right"
  )

# save plot
ggsave(str_c(outfolder,"Overview-plots/",date,"_Correlation_plots_DNA-vs-PCR-vs-Reads.png"), width = 22, height = 10, units = "cm")
ggsave(str_c(outfolder,"Overview-plots/",date,"_Correlation_plots_DNA-vs-PCR-vs-Reads.pdf"), width = 22, height = 10, units = "cm") 
```

# Part 2: Global analysis
## Tidying & cleaning data for plotting
Within the next lines of code, raw data is tidied up and prepared for plotting. Full dataset containing all reads obtained as well as downsampled datasets are processed. 

### Full dataset
Data from April:
```{r eval=FALSE}
# read in family data
classified_reads_april_family <- read_tsv(destfile_april_family)
classified_reads_april_family <- rename(classified_reads_april_family, Name = X1) %>%
    mutate(TaxRank = "F")

# read in genus data
classified_reads_april_genus <- read_tsv(destfile_april_genus)
classified_reads_april_genus <- rename(classified_reads_april_genus, Name = X1) %>%
    mutate(TaxRank = "G")

# combine family and genus data
classified_reads_april <- rbind(classified_reads_april_family, classified_reads_april_genus)

# tidy data and extract absolute read numbers
tidy_classified_reads_april <- gather(classified_reads_april, key = "Location", value = "Reads", 2:13)%>%
  mutate(Time = "April") %>%
  mutate(Classifyer = classifyr)

# calculate percentage
total_read_number_april_family <- 
  hash(keys=colnames(classified_reads_april_family[2:13]),
  values=colSums(classified_reads_april_family[2:13]))
total_read_number_april_genus <- 
  hash(keys=colnames(classified_reads_april_genus[2:13]),
  values=colSums(classified_reads_april_genus[2:13]))

for (i in 1:nrow(tidy_classified_reads_april)){
  if(tidy_classified_reads_april$TaxRank[i] == "F"){
    tidy_classified_reads_april$Percentage[i] <- tidy_classified_reads_april$Reads[i]/total_read_number_april_family [[tidy_classified_reads_april$Location[i]]]*100
  } else {
    tidy_classified_reads_april$Percentage[i] <- tidy_classified_reads_april$Reads[i]/total_read_number_april_genus [[tidy_classified_reads_april$Location[i]]]*100
  } 
} 

# clean and order dataset
clean_classified_reads_april <- select(tidy_classified_reads_april, Name, TaxRank, Location, Time, Classifyer, Reads, Percentage) %>%
  mutate_at(vars(TaxRank,Location,Time,Classifyer),list(factor))

clean_classified_reads_april # display cleaned data table
```

Data from June:
```{r eval=FALSE}
# read in family data
classified_reads_june_family <- read_tsv(destfile_june_family)
classified_reads_june_family <- rename(classified_reads_june_family, Name = X1) %>%
    mutate(TaxRank = "F")

# read in genus data
classified_reads_june_genus <- read_tsv(destfile_june_genus)
classified_reads_june_genus <- rename(classified_reads_june_genus, Name = X1) %>%
    mutate(TaxRank = "G")

# combine family and genus data
classified_reads_june <- rbind(classified_reads_june_family, classified_reads_june_genus)

# tidy data and extract absolute read numbers
tidy_classified_reads_june <- gather(classified_reads_june, key = "Location", value = "Reads", 2:13)%>%
  mutate(Time = "June") %>%
  mutate(Classifyer = classifyr)

# calculate percentage
total_read_number_june_family <- 
  hash(keys=colnames(classified_reads_june_family[2:13]),
  values=colSums(classified_reads_june_family[2:13]))
total_read_number_june_genus <- 
  hash(keys=colnames(classified_reads_june_genus[2:13]),
  values=colSums(classified_reads_june_genus[2:13]))

for (i in 1:nrow(tidy_classified_reads_june)){
  if(tidy_classified_reads_june$TaxRank[i] == "F"){
    tidy_classified_reads_june$Percentage[i] <- tidy_classified_reads_june$Reads[i]/total_read_number_june_family [[tidy_classified_reads_june$Location[i]]]*100
  } else {
    tidy_classified_reads_june$Percentage[i] <- tidy_classified_reads_june$Reads[i]/total_read_number_june_genus [[tidy_classified_reads_june$Location[i]]]*100
  } 
} 

# clean and order dataset
clean_classified_reads_june <- select(tidy_classified_reads_june, Name, TaxRank, Location, Time, Classifyer, Reads, Percentage) %>%
  mutate_at(vars(TaxRank,Location,Time,Classifyer),list(factor))

clean_classified_reads_june # display cleaned data table
```
Data from August:
```{r eval=FALSE}
# read in family data
classified_reads_august_family <- read_tsv(destfile_august_family)
classified_reads_august_family <- rename(classified_reads_august_family, Name = X1) %>%
    mutate(TaxRank = "F")

# read in genus data
classified_reads_august_genus <- read_tsv(destfile_august_genus)
classified_reads_august_genus <- rename(classified_reads_august_genus, Name = X1) %>%
    mutate(TaxRank = "G")

# combine family and genus data
classified_reads_august <- rbind(classified_reads_august_family, classified_reads_august_genus)

# tidy data and extract absolute read numbers
tidy_classified_reads_august <- gather(classified_reads_august, key = "Location", value = "Reads", 2:13)%>%
  mutate(Time = "August") %>%
  mutate(Classifyer = classifyr)

# calculate percentage
total_read_number_august_family <- 
  hash(keys=colnames(classified_reads_august_family[2:13]),
  values=colSums(classified_reads_august_family[2:13]))
total_read_number_august_genus <- 
  hash(keys=colnames(classified_reads_august_genus[2:13]),
  values=colSums(classified_reads_august_genus[2:13]))

for (i in 1:nrow(tidy_classified_reads_august)){
  if(tidy_classified_reads_august$TaxRank[i] == "F"){
    tidy_classified_reads_august$Percentage[i] <- tidy_classified_reads_august$Reads[i]/total_read_number_august_family [[tidy_classified_reads_august$Location[i]]]*100
  } else {
    tidy_classified_reads_august$Percentage[i] <- tidy_classified_reads_august$Reads[i]/total_read_number_august_genus [[tidy_classified_reads_august$Location[i]]]*100
  } 
} 

# clean and order dataset
clean_classified_reads_august <- select(tidy_classified_reads_august, Name, TaxRank, Location, Time, Classifyer, Reads, Percentage) %>%
  mutate_at(vars(TaxRank,Location,Time,Classifyer),list(factor))

clean_classified_reads_august # display cleaned data table
```
Combining data of all three time points:
```{r eval=FALSE}
clean_classified_reads <- bind_rows(clean_classified_reads_april, clean_classified_reads_june, clean_classified_reads_august)%>%
  transform( Time = factor(Time, levels = c("April","June","August")))

classified_reads_root <- read_tsv(destfile_root)
clean_classified_reads <-rbind(clean_classified_reads, classified_reads_root)

clean_classified_reads # display combined data table

```

### Downsampled data
Similar to the full datasets, the sets of downsampled data are prepared for plotting. 
```{r eval=FALSE}

# Initialise lists
list1 <- list("","","","")
list2 <- list("","","","")
tidy_classified_reads_downsampled <- list(list1,list2)
clean_classified_reads_downsampled <- list(list1,list2)

# loop over all 4 downsampling sets and bot family and genus levels
for (set in 1:4){
  for (level in 1:2){
    if(level == 1){
      # create hashs of taxon classification and associated number of reads
      total_read_number <- hash(keys=colnames(classified_reads_downsampled[[level]][[set]][2:21]), values=colSums(classified_reads_downsampled[[level]][[set]][2:21]))
      # convert dataframe so that factors can be used for plotting    
      tidy_classified_reads_downsampled[[level]][[set]] <- gather(classified_reads_downsampled[[level]][[set]], key = "Location", value = "Reads", 2:21)
    }  else {
      # create hashs of taxon classification and associated number of reads
      total_read_number <- hash(keys=colnames(classified_reads_downsampled[[level]][[set]][2:19]), values=colSums(classified_reads_downsampled[[level]][[set]][2:19]))
      # convert dataframe so that factors can be used for plotting
      tidy_classified_reads_downsampled[[level]][[set]] <- gather(classified_reads_downsampled[[level]][[set]], key = "Location", value = "Reads", 2:19)
    }
   
    # calculate and add percentage
    for (i in 1:nrow(tidy_classified_reads_downsampled[[level]][[set]])){
      tidy_classified_reads_downsampled[[level]][[set]]$Percentage[i] <- tidy_classified_reads_downsampled[[level]][[set]]$Reads[i]/total_read_number [[tidy_classified_reads_downsampled[[level]][[set]]$Location[i]]]*100
    }
    
    # clean dataframe
    clean_classified_reads_downsampled[[level]][[set]] <- separate(tidy_classified_reads_downsampled[[level]][[set]], Location, into = c("Location","Time"), sep="-") %>%
      separate(Time, into = c("Time","TaxRank"), sep="_") %>%
      mutate(TaxRank = str_replace(TaxRank, pattern = "family.tsv", "F")) %>%
      mutate(TaxRank = str_replace(TaxRank, pattern = "genus.tsv", "G")) %>%
      mutate(Info = str_c("downsampled ",classifyr)) %>%
      mutate_at(vars(TaxRank,Location,Time,Info),funs(factor))
    
    colnames(clean_classified_reads_downsampled[[level]][[set]])[colnames(clean_classified_reads_downsampled[[level]][[set]]) == "X1"] <- "Name"
  }  
}
```

## Plotting data
### Reads per location/time point
Next, the distribution of reads per location/time point is displayed in a barchart.
```{r}
# Create output directory
system(str_c("mkdir ",outfolder,"Overview-plots/"))

subset1<- filter(clean_classified_reads, Name == "root" | Name == "unclassified") %>%
  mutate( Name = str_replace(Name, pattern = "root", "classified")) %>%
  transform( Name = factor(Name, levels = c("unclassified","classified")))

# Overview plot: total number of reads per location
ggplot(data = subset1, mapping = aes(x = Location, y = Reads, fill = Time))+
  geom_bar(stat="identity", width = 0.7, position ="dodge", colour = "black")+
  labs(title = "Number of reads per location/time point")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5, face="bold.italic"))+
  theme(legend.title = element_blank())+
  scale_y_continuous(labels = scientific_1)+
  scale_fill_manual(values = tcolors) +
  theme(axis.line.x = element_blank())+
  theme(axis.ticks.x = element_blank())

# save plot
ggsave(str_c(outfolder,"Overview-plots/",date,"_Number-of-reads-per-location-time-point_",classifyr,".png"), width = 20, height = 15, units = "cm")
ggsave(str_c(outfolder,"Overview-plots/",date,"_Number-of-reads-per-location-time-point_",classifyr,".pdf"), width = 20, height = 15, units = "cm")
```

### Number of classified/unclassified reads per location/time point
Next, the distribution of classified/unclassified reads per location/time point is displayed.
```{r}
# create a subset of the dataframe for plotting
subset2<- filter(clean_classified_reads, Name == "root" | Name == "unclassified") %>%
  mutate( Name = str_replace(Name, pattern = "root", "classified")) %>%
  transform( Name = factor(Name, levels = c("unclassified","classified"))) %>%
  mutate( Location = str_replace(Location, pattern = "Mock", "P"))

# Overview: Distribution of classified vs. unclassified reads per location and time point
ggplot(data = subset2, mapping = aes(x = Location, y = Reads, fill = Name))+
  geom_bar(stat="identity", width = 0.7)+
  facet_wrap(~Time)+
  geom_hline(aes(yintercept=37000, color="Downsampling"), linetype="dashed")+
  labs(title = "Distribution of classified vs. unclassified reads per location and time point")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = 11))+
  theme(axis.text.x = element_text(size = 7))+
  theme(axis.text.y = element_text(size = 7))+
  theme(legend.text = element_text(size = 8))+
  theme(legend.title = element_blank())+
  scale_fill_manual(values=c("#999999","#56B4E9"))+
  scale_y_continuous(labels = scientific_1)

ggsave(str_c(outfolder,"Overview-plots/",date,"_Distribution-of-classified-unclassified-reads-per-location-time-point_V1_",classifyr,".png"), width = 20, height = 15, units = "cm")
ggsave(str_c(outfolder,"Overview-plots/",date,"_Distribution-of-classified-unclassified-reads-per-location-time-point_V1_",classifyr,".pdf"), width = 20, height = 15, units = "cm") 

# Zoom in: Distribution of classified vs. unclassified reads per location and time point
ggplot(data = subset2, mapping = aes(x = Location, y = Reads, fill = Name))+
  geom_bar(stat="identity", width = 0.7)+
  facet_wrap(~Time)+
  geom_hline(aes(yintercept=37000, color="Downsampling"), linetype="dashed")+
  labs(title = "Zoom in: Distribution of classified vs. unclassified reads per location and time point")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = 11))+
  theme(axis.text.x = element_text(size = 7))+
  theme(axis.text.y = element_text(size = 7))+
  theme(legend.text = element_text(size = 8))+
  theme(legend.title = element_blank())+
  scale_fill_manual(values=c("#999999","#56B4E9"))+
  coord_cartesian(ylim = c(0,80000))+
  scale_y_continuous(labels = scientific_1)

# save plot
ggsave(str_c(outfolder,"Overview-plots/",date,"_Distribution-of-classified-unclassified-reads-per-location-time-point_V2_",classifyr,".png"), width = 20, height = 15, units = "cm")
ggsave(str_c(outfolder,"Overview-plots/",date,"_Distribution-of-classified-unclassified-reads-per-location-time-point_V2_",classifyr,".pdf"), width = 20, height = 15, units = "cm") 

# Distribution of normalized numbers of classified/unclassified reads per sample location
ggplot(data = subset2, mapping = aes(x = Location, y = Percentage, fill = Name))+
  geom_bar(stat="identity", width = 0.7)+
  facet_wrap(~Time)+  
  labs(title = "Percentage of classified vs. unclassified reads per location and time point", y = "% Reads")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = 11))+
  theme(axis.text.x = element_text(size = 8))+
  theme(axis.text.y = element_text(size = 7))+
  theme(legend.text = element_text(size = 8))+
  theme(legend.title = element_blank())+
  scale_fill_manual(values=c("#999999","#56B4E9"))

# save plot
ggsave(str_c(outfolder,"Overview-plots/",date,"_Percentage-of-classified-unclassified-reads-per-location-time-point_",classifyr,".png"), width = 20, height = 15, units = "cm")
ggsave(str_c(outfolder,"Overview-plots/",date,"_Percentage-of-classified-unclassified-reads-per-location-time-point_",classifyr,".pdf"), width = 20, height = 15, units = "cm") 
```

### Number of classified reads per taxonomic level (Supplementary Figure)
The number of reads classified onto the different taxonomic levels was displayed using the followiung code:
```{r}
# Filtering and adjutsing data for plotting
subset3_1 <- filter(clean_classified_reads) %>% 
  mutate(TaxRank = as.character(TaxRank)) %>%
  mutate(TaxRank = ifelse(Name == "root", "R", TaxRank)) %>%
  mutate(TaxRank = factor(TaxRank, levels = c("U","R","D","P","C","O","F","G","K","-")))

subset3_2 <- subset3_1 %>%
  filter(TaxRank != "-" & TaxRank != "K" ) %>%
  filter(!is.na(Reads)) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "C", "Class")) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "R", "Classified")) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "U", "Unclassified")) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "D", "Domain")) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "P", "Phylum")) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "O", "Order")) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "F", "Family")) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "G", "Genus")) %>%
  group_by(TaxRank, Location, Time) %>%
  summarize(Sum = sum(Reads))%>%
  ungroup(TaxRank, Location, Time)

subset3_3 <- filter(subset3_1, TaxRank == "R" | TaxRank == "U" ) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "R", "Classified")) %>%
  mutate( TaxRank = str_replace(TaxRank, pattern = "U", "Unclassified")) %>%
  filter(!is.na(Reads)) %>%
  group_by(Location, Time) %>%
  summarize(Sum = sum(Reads)) %>%
  mutate(TaxRank = "Total")

subset3_4 <- full_join(subset3_2, subset3_3, by = c("TaxRank", "Location", "Time", "Sum"))

subset3_4 <- filter(subset3_4, TaxRank != "Unclassified") %>%
  mutate(Location = factor(Location)) %>%
  mutate(TaxRank = factor(TaxRank, levels = c("Total","Classified","Domain","Phylum","Class","Order","Family","Genus"), ordered = TRUE))

for (setting in c("samples","samples+controls")){
  if (setting == "samples"){
    subset3_5 <- filter(subset3_4, Location != "N" & Location != "P")
  } else {
    subset3_5 <- mutate(subset3_4, Location = str_replace(Location, pattern = "Mock", "P"))
  }
  
  # Number of classified reads per taxonomic level
  ggplot(data = subset3_5, mapping = aes(x = Location, y = Sum, fill = TaxRank))+
    geom_bar(stat="identity",position = position_dodge(), width = 0.7)+
    facet_grid(cols = vars(Time))+ 
    geom_hline(yintercept=37000, linetype="dashed", color = "black")+
    labs(title = "Number of classified reads per taxonomic level",y = "Reads (1)")+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5, size = 22))+
    theme(axis.text.x = element_text(size = 16))+
    theme(axis.text.y = element_text(size = 14))+
    theme(legend.text = element_text(size = 16))+
    theme(legend.title = element_blank())+
    scale_fill_grey()+
    scale_y_continuous(trans='log10', breaks = trans_breaks("log10", function(x) 10^x),
                       labels = trans_format("log10", math_format(10^.x)))+
    annotation_logticks(sides="l")
  
  # save plots
  ggsave(str_c(outfolder,"Overview-plots/",date,"_Number-of-classified-reads-per-taxonomic-level_",classifyr,"_",setting,".png"), width = 30, height = 15, units = "cm")
  ggsave(str_c(outfolder,"Overview-plots/",date,"_Number-of-classified-reads-per-taxonomic-level_",classifyr,"_",setting,".pdf"), width = 30, height = 15, units = "cm") 
}
```

### Relative abundance: Representation on the different taxonomic levels
The relative species abundance on the different taxonomic levels was visualised using the following lines of code. 
#### Full dataset
```{r}
# create output directory
system(str_c("mkdir ",outfolder,"Relatve_abundance"))
system(str_c("mkdir ",outfolder,"Relatve_abundance/Full-dataset/"))
system(str_c("mkdir ",outfolder,"Relatve_abundance/Full-dataset/Total"))
system(str_c("mkdir ",outfolder,"Relatve_abundance/Full-dataset/Top10-per-sample"))
system(str_c("mkdir ",outfolder,"Relatve_abundance/Full-dataset/Top10-highest-single-abundance"))
system(str_c("mkdir ",outfolder,"Relatve_abundance/Full-dataset/Top10-highest-total-abundance"))

# define color vector for plots
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
col_vector_3 = c(col_vector,col_vector,col_vector)
#pie(rep(1,60), col=sample(col_vector, 60)) # to visualise color vector

# make plots for all samples as well as samples+controls
for (setting in c("samples","samples+controls")){
  
  # loop over all taxonomic ranks
  ranks <- c("family","genus")
  ranks_plural <- c("families","genera")
  rank_i <- 0
  
  for (rank in c("F","G")){
    rank_i <- rank_i + 1
    subset_rank<- filter(clean_classified_reads, TaxRank == rank, Percentage != "NA")%>%  
      mutate( Location = str_replace(Location, pattern = "Mock", "P")) %>%
      arrange(desc(Percentage))
    
    # filter subset of data so that only the samples or samples + controls are plotted
    if (setting == "samples"){
      subset_rank_1 <- filter(subset_rank, Location != "N" & Location != "P")
    } else {
      subset_rank_1 <- subset_rank
    }
    
    # save selection
    write_tsv(subset_rank_1, str_c(outfolder,"Relatve_abundance/Full-dataset/Total/",date,"_subset_rank_1-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,".txt"), col_names = TRUE)
    
    # Where we at?
    print(paste0('Processing data of ', setting,' - on ', ranks[rank_i], ' level'))
    
    # a. plot relativ adunance showing all taxa for ranks = "D" || "P" || "C"
    if (rank == "D" || rank == "P" || rank == "C" ) {
      
      ggplot(data = subset_rank_1, mapping = aes(x = Location, y = Percentage, fill = Name))+
          geom_bar(stat="identity",  position = "fill", width = 0.8)+
          facet_wrap(~Time)+  
          labs(title = str_c("Relativ abundance on ",ranks[rank_i]," level"), y = "Abundance")+
          theme_bw()+
          theme(plot.title = element_text(hjust = 0.5, size = 11))+
          theme(axis.text.x = element_text(size = 8))+
          theme(axis.text.y = element_text(size = 7))+
          theme(legend.text = element_text(size = 8))+
          theme(legend.title = element_blank())+
          scale_fill_manual(values = col_vector_3)
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Total/",date,"_Relative-abundance-plot_Total_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,".pdf"), width = 50, height = 20, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Total/",date,"_Relative-abundance-plot_Total_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,".png"), width = 50, height = 20, units = "cm")
    } 
    
    # for all taxa for ranks != "D" 
    if (rank != "D") {
      
      # b. select top 10 most abundant taxa per sample
      subset_rank_top10<- subset_rank_1 %>%
        arrange(Time) %>%
        arrange(Location) %>%
        unite_("Location_Time", c("Location","Time"))%>%
        group_by(Location_Time)
      subset_rank_top10_top <- slice(subset_rank_top10, -11:-n())
      subset_rank_top10_bottom <- slice(subset_rank_top10, 11:n()) %>%
        mutate( Name = "Other")
      subset_rank_top10 <- bind_rows(subset_rank_top10_top, subset_rank_top10_bottom)%>%
        separate(Location_Time, into = c("Location","Time"), sep="_")%>%
        transform( Time = factor(Time, levels = c("April","June","August"))) 
      
      legend <- subset_rank_top10$Name %>%
        unique()
      legend <- fct_relevel(legend, "Other", after = 0L)
      subset_rank_top10 <- transform(subset_rank_top10, Name = factor(Name, levels = legend))
      
      # plot relativ abundance showing the top 10 most abundant taxa per sample for ranks != "D"
      ggplot(data = subset_rank_top10, mapping = aes(x = Location, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Time)+  
        labs(title = str_c("Top 10 most abundant ",ranks_plural[rank_i]," per sample"), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = col_vector)
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-per-sample/",date,"_Relative-abundance-plot_Top10_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,".pdf"), width = 30, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-per-sample/",date,"_Relative-abundance-plot_Top10_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,".png"), width = 30, height = 15, units = "cm")
      
      
      # c. select top 10 most abundant taxa with the highest single abundance (in percent) among all samples
      legend <- subset_rank_1$Name %>%
        unique()
      legend <- c(legend[1:10], "Other")
      subset_rank_top10_2 <- transform(subset_rank_1, Name = factor(Name, levels = legend)) %>%
        mutate(Name = replace_na(Name, "Other")) %>%
        transform( Time = factor(Time, levels = c("April","June","August"))) 
    
      # plot relativ abundance (Version 1) showing the top 10 most abundant taxa with the highest single abundance (in percent) among all samples
      ggplot(data = subset_rank_top10_2, mapping = aes(x = Location, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Time)+  
        labs(title = str_c("Most abundant ",ranks_plural[rank_i]), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V1.pdf"), width = 30, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V1.png"), width = 30, height = 15, units = "cm")
    
      # plot relativ abundance (Version 2) showing the top 10 most abundant taxa with the highest single abundance (in percent) among all samples
      ggplot(data = subset_rank_top10_2, mapping = aes(x = Time, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Location)+  
        labs(title = str_c("Most abundant ",ranks_plural[rank_i]), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V2.pdf"), width = 30, height = 25, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V2.png"), width = 30, height = 25, units = "cm")
     
      
      # d. select top 10 most abundant taxa with the highest total abundance (percentage) among all samples
       subset_tmp <- subset_rank_1 %>%
        group_by(Name) %>%
        summarize(Sum = sum(Percentage)) %>%
        ungroup() %>%
        arrange(desc(Sum))
      legend <- subset_tmp$Name %>%
        unique()
      legend <- c(legend[1:10], "Other")
      subset_rank_top10_3 <- transform(subset_rank_1, Name = factor(Name, levels = legend)) %>%
        mutate(Name = replace_na(Name, "Other")) %>%
        transform( Time = factor(Time, levels = c("April","June","August"))) 
    
      # plot relativ abundance (Version 1) showing the top 10 most abundant taxa with the highest total abundance (percentage) among all samples
      ggplot(data = subset_rank_top10_3, mapping = aes(x = Location, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Time)+  
        labs(title = str_c("Most abundant ",ranks_plural[rank_i]), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V1.pdf"), width = 30, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V1.png"), width = 30, height = 15, units = "cm")
    
      # plot relativ abundance (Version 2) showing the top 10 most abundant taxa with the highest total abundance (percentage) among all samples
      ggplot(data = subset_rank_top10_3, mapping = aes(x = Time, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Location)+  
        labs(title = str_c("Most abundant ",ranks_plural[rank_i]), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V2.pdf"), width = 30, height = 25, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V2.png"), width = 30, height = 25, units = "cm")
      
      # shorten annotation for August
      subset_rank_top10_3_1 <- subset_rank_top10_3 %>%
        mutate(Time = str_replace(Time, pattern = "August", "Aug")) %>%
        transform(Time = factor(Time, levels = c("April","June","Aug"), ordered = TRUE))
      
      # plot relativ abundance (Version 3) the top 10 most abundant taxa with the highest total abundance (percentage) among all samples
      ggplot(data = subset_rank_top10_3_1, mapping = aes(x = Time, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity", position = "fill", width = 0.4, alpha=0.5)+
        facet_grid(cols = vars(Location))+ 
        labs(title = str_c("Most abundant ",ranks_plural), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 22))+
        theme(axis.text.x = element_text(size = 16))+
        theme(axis.text.y = element_text(size = 14))+
        theme(legend.text = element_text(size = 16))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plot
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V3.pdf"), width = 60, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Full-dataset/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks[rank_i],"_",classifyr,"_",setting,"_V3.png"), width = 60, height = 15, units = "cm")
      
    }
  }
}
```

#### Downsampled data
```{r}
# create general output directories
system(str_c("mkdir ",outfolder,"Relatve_abundance"))
system(str_c("mkdir ",outfolder,"Relatve_abundance/Downsampled-dataset/"))
system(str_c("mkdir ",outfolder,"Relatve_abundance/Downsampled-dataset/family"))
system(str_c("mkdir ",outfolder,"Relatve_abundance/Downsampled-dataset/genus"))

# for each of the downsampling datasets poduced 
for (set in 1:4){
  for (level in 1:2){
  
    if (level == 1){
      rank <- "F"                 
      rank_i <- 5                 
      ranks <- "Family"            
      ranks_plural <- "families"     
    } else {
      rank <- "G"                  
      rank_i <- 6                 
      ranks <- "Genus"            
      ranks_plural <- "genera"     
    }
    
    # create specific output directories
    system(str_c("mkdir ",outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set))
    system(str_c("mkdir ",outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-per-sample"))
    system(str_c("mkdir ",outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-single-abundance"))
    system(str_c("mkdir ",outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance"))
    
    # define color vector for plots
    qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
    col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
    col_vector_3 = c(col_vector,col_vector,col_vector)
    #pie(rep(1,60), col=sample(col_vector, 60)) # to visualise color vector
    
    # make plots for all samples as well as samples+controls
    for (setting in c("samples","samples+controls")){
      
      # Where we at?
      print(paste0('Processing downsampled data set ',set,' on ', ranks,' level, for ', setting))
      
      subset_rank<- filter(clean_classified_reads_downsampled[[level]][[set]], TaxRank == rank, Percentage != "NA")%>%  
        #filter(!grepl('Eukaryota', Linage)) %>% # select only bacterial taxa by filterin out eukaryotic taxa below domain level
        #filter(!grepl('Archaea', Linage)) %>% # select only bacterial taxa by filterin out archaea taxa below domain level
        mutate( Location = str_replace(Location, pattern = "Mock", "P")) %>%
        arrange(desc(Percentage))
        
      # filter subset of data so that only the samples or samples + controls are plotted
      if (setting == "samples"){
        subset_rank_1 <- filter(subset_rank, Location != "N" & Location != "P")
      } else {
        subset_rank_1 <- subset_rank
      }
          
      # a. select top 10 most abundant taxa per sample
      subset_rank_top10<- subset_rank_1 %>% 
        arrange(Time) %>%
        arrange(Location) %>%
        unite_("Location_Time", c("Location","Time"))%>%
        group_by(Location_Time)
      subset_rank_top10_top <- slice(subset_rank_top10, -11:-n())
      subset_rank_top10_bottom <- slice(subset_rank_top10, 11:n()) %>%
        mutate( Name = "Other")
      subset_rank_top10 <- bind_rows(subset_rank_top10_top, subset_rank_top10_bottom)%>%
        separate(Location_Time, into = c("Location","Time"), sep="_")%>%
        transform( Time = factor(Time, levels = c("April","June","August"))) 
        
      legend <- subset_rank_top10$Name %>%
        unique()
      legend <- fct_relevel(legend, "Other", after = 0L)
      subset_rank_top10 <- transform(subset_rank_top10, Name = factor(Name, levels = legend))
        
      # plot relativ abundance showing the top 10 most abundant taxa per sample for ranks != "D"
      ggplot(data = subset_rank_top10, mapping = aes(x = Location, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Time)+  
        labs(title = str_c("Top 10 most abundant ",ranks_plural," per sample"), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = col_vector)
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-per-sample/",date,"_Relative-abundance-plot_Top10_level",rank_i,"_",ranks,"_",classifyr,"_",setting,".pdf"), width = 30, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-per-sample/",date,"_Relative-abundance-plot_Top10_level",rank_i,"_",ranks,"_",classifyr,"_",setting,".png"), width = 30, height = 15, units = "cm")
          
          
      # b. select top 10 most abundant taxa with the highest single abundance (in percent) among all samples
      legend <- subset_rank_1$Name %>%
        unique()
      legend <- c(legend[1:10], "Other")
      subset_rank_top10_2 <- transform(subset_rank_1, Name = factor(Name, levels = legend)) %>%
          mutate(Name = replace_na(Name, "Other")) %>%
          transform( Time = factor(Time, levels = c("April","June","August"))) 
    
      # plot relativ abundance (Version 1) showing the top 10 most abundant taxa with the highest single abundance (in percent) among all samples
      ggplot(data = subset_rank_top10_2, mapping = aes(x = Location, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Time)+  
        labs(title = str_c("Most abundant ",ranks_plural), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V1.pdf"), width = 30, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V1.png"), width = 30, height = 15, units = "cm")
      
      # plot relativ abundance (Version 2) showing the top 10 most abundant taxa with the highest single abundance (in percent) among all samples
      ggplot(data = subset_rank_top10_2, mapping = aes(x = Time, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Location)+  
        labs(title = str_c("Most abundant ",ranks_plural), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V2.pdf"), width = 30, height = 25, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V2.png"), width = 30, height = 25, units = "cm")
        
      # shorten annotation for August
      subset_rank_top10_2_1 <- subset_rank_top10_2 %>%
        mutate(Time = str_replace(Time, pattern = "August", "Aug")) %>%
        transform(Time = factor(Time, levels = c("April","June","Aug"), ordered = TRUE))
      
      # plot relativ abundance (Version 3) showing the top 10 most abundant taxa with the highest single abundance (in percent) among all samples
      ggplot(data = subset_rank_top10_2_1, mapping = aes(x = Time, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity", position = "fill", width = 0.4)+
        facet_grid(cols = vars(Location))+ 
        labs(title = str_c("Most abundant ",ranks_plural), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 22))+
        theme(axis.text.x = element_text(size = 16))+
        theme(axis.text.y = element_text(size = 14))+
        theme(legend.text = element_text(size = 16))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plot
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V3.pdf"), width = 60, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-single-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V3.png"), width = 60, height = 15, units = "cm")
        
      
      # c. select top 10 most abundant taxa with the highest total abundance (percentage) among all samples
      subset_tmp <- subset_rank_1 %>%
        group_by(Name) %>%
        summarize(Sum = sum(Percentage)) %>%
        ungroup() %>%
        arrange(desc(Sum))
      legend <- subset_tmp$Name %>%
        unique()
      legend <- c(legend[1:10], "Other")
      subset_rank_top10_3 <- transform(subset_rank_1, Name = factor(Name, levels = legend)) %>%
        mutate(Name = replace_na(Name, "Other")) %>%
        transform( Time = factor(Time, levels = c("April","June","August"))) 
      
      # save selection
      write(legend, file =str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_legend_",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V3.txt"))
      
      # plot relativ abundance (Version 1) showing the top 10 most abundant taxa with the highest total abundance (percentage) among all samples
      ggplot(data = subset_rank_top10_3, mapping = aes(x = Location, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Time)+  
        labs(title = str_c("Most abundant ",ranks_plural), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V1.pdf"), width = 30, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V1.png"), width = 30, height = 15, units = "cm")
        
      # plot relativ abundance (Version 2) showing the top 10 most abundant taxa with the highest total abundance (percentage) among all samples
      ggplot(data = subset_rank_top10_3, mapping = aes(x = Time, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity",  position = "fill", width = 0.8)+
        facet_wrap(~Location)+  
        labs(title = str_c("Most abundant ",ranks_plural), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plots
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V2.pdf"), width = 30, height = 25, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Total-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V2.png"), width = 30, height = 25, units = "cm")
      
      # shorten annotation for August
      subset_rank_top10_3_1 <- subset_rank_top10_3 %>%
        mutate(Time = str_replace(Time, pattern = "August", "Aug")) %>%
        transform(Time = factor(Time, levels = c("April","June","Aug"), ordered = TRUE))
      
      # save selection
      write_tsv(subset_rank_top10_3_1, str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_subset_rank_top10_3_1_",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V3.txt"), col_names = TRUE)
      
      # plot relativ abundance (Version 3) the top 10 most abundant taxa with the highest total abundance (percentage) among all samples
      ggplot(data = subset_rank_top10_3_1, mapping = aes(x = Time, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity", position = "fill", width = 0.4)+
        facet_grid(cols = vars(Location))+ 
        labs(title = str_c("Most abundant ",ranks_plural), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 22))+
        theme(axis.text.x = element_text(size = 16))+
        theme(axis.text.y = element_text(size = 14))+
        theme(legend.text = element_text(size = 16))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plot
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V3.pdf"), width = 60, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,"_V3.png"), width = 60, height = 15, units = "cm")
      
      
      # for the first downsampling set
      if (set == 1){
        
        # extract information about total number of reads assigned to the top 10 most abundant families
        subset_rank_info <- subset(subset_rank_top10_3, select = c(Name, Location, Time, Reads)) %>%
          mutate(Name = ifelse(Name != "Other", str_c("10 most abundant ",ranks_plural), str_c("Other ",ranks_plural)))%>%
          group_by(Name, Location, Time) %>%
          summarize(Sum = sum(Reads)) %>%
          ungroup(Name, Location, Time) 
        
        # use helper dataframes to set all unexiting values in subset_rank_info to 0
        help_subset_1 <- subset(subset3_4, select = c(Location, Time)) %>%
          group_by(Location, Time) %>%
          unique()%>%
          mutate(Name = str_c("10 most abundant ",ranks_plural))
        help_subset_2 <- subset(subset3_4, select = c(Location, Time)) %>%
          group_by(Location, Time) %>%
          unique()%>%
          mutate(Name = str_c("Other ",ranks_plural))
        help_subset <- rbind(help_subset_1 , help_subset_2) %>%
          ungroup(Location, Time) %>%
          mutate(Location = as.character(Location)) %>%
          mutate( Location = str_replace(Location, pattern = "Mock", "P"))
        
        # join helper dataframe with info dataframe to get NA values
        subset_rank_info <- right_join(subset_rank_info, help_subset) 
        
        # extract information about number of total reads, number of classified reads & number of on family level classified reads
        subset_rank_2 <- subset(subset3_4, select = c(TaxRank, Location, Time, Sum)) %>%
          filter(TaxRank == "Total" | TaxRank == "Classified" | TaxRank == ranks) %>%
          rename(Name = TaxRank) %>%
          mutate( Location = str_replace(Location, pattern = "Mock", "P"))
        
        # combine both dataframes and order factors for plotting
        subset_rank_3 <- rbind(subset_rank_2, subset_rank_info) %>%
          transform(Name = factor(Name, levels = c("Total","Classified",ranks,str_c("10 most abundant ",ranks_plural),str_c("Other ",ranks_plural)), ordered = TRUE)) %>%
          mutate(Time = str_replace(Time, pattern = "August", "Aug")) %>%
          transform(Time = factor(Time, levels = c("April","June","Aug"), ordered = TRUE))
        
        # filter subset of data so that only the samples or samples + controls are plotted
        if (setting == "samples"){
          subset_rank_3  <- filter(subset_rank_3 , Location != "N" & Location != "P")
        }
        
        # plot number of reads for the different levels (Total, classified, family/genus, 10 most abundant families/genera, Other families/genera) 
        ggplot(data = subset_rank_3, mapping = aes(x = Time, y = Sum, fill = Name))+
          geom_bar(stat="identity", position = position_dodge(), width = 0.7)+
          facet_grid(cols = vars(Location))+ 
          geom_hline(yintercept=37000, linetype="dashed", color = "black")+
          labs(title = "Level of Classifiation",y = "Reads (1)")+
          theme_bw()+
          theme(plot.title = element_text(hjust = 0.5, size = 22))+
          theme(axis.text.x = element_text(size = 16))+
          theme(axis.text.y = element_text(size = 14))+
          theme(legend.text = element_text(size = 16))+
          theme(legend.title = element_blank())+
          scale_fill_manual(values=c("#636363","#bdbdbd","#d9d9d9","#56B4E9","#ccebc5"))+
          scale_y_continuous(trans='log10', breaks = trans_breaks("log10", function(x) 10^x),labels = trans_format("log10", math_format(10^.x)))+
          annotation_logticks(sides="l")
        # save plot
        ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_Read-numbers-in-comparison_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,".pdf"), width = 60, height = 15, units = "cm")
        ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/",ranks,"/Set",set,"/Top10-highest-total-abundance/",date,"_Read-numbers-in-comparison_level-",rank_i,"_",ranks,"_",classifyr,"_",setting,".png"), width = 60, height = 15, units = "cm") 
      }
    }
  }
}
```
#### Extra: Main Figure 
Overlay of Relative abundance plots from Downsampled data with full dataset.
Plot relativ abundance (Version 4) of the top 10 most abundant taxa with the highest total abundance (percentage) among all downsampled samples (set1) and add corresponding abundances for samples with <37,000 but >1,000 reads from the full dataset
```{r}
# load legend of top 10 most abundant
legend <- read_tsv(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/family/Set1/Top10-highest-total-abundance/",date,"_legend_5_family_",classifyr,"_samples_V3.txt"), col_names = FALSE)
legend <- legend$X1
# load top 10 list from downsamled data
subset_downsampling <- read_tsv(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/family/Set1/Top10-highest-total-abundance/",date,"_subset_rank_top10_3_1_5_family_",classifyr,"_samples_V3.txt"), col_names = TRUE)
# load relative abundances based on full dataset 
subset_full <- read_tsv(str_c(outfolder,"Relatve_abundance/Full-dataset/Total/",date,"_subset_rank_1-1_family_",classifyr,"_samples.txt"), col_names = TRUE)

# combine both datasets and 
subset_full_1 <- select(subset_full, -Classifyer)
subset_full_1$Name <- as.factor(subset_full_1$Name)
subset_full_1$Info <- "full dataset minimap"

# define which locations are already existing in the downsampled data
locations <- subset_downsampling %>%
        unite("Colnames", c("Location","Time"), sep = " - ") %>%
        select("Colnames") %>%
        unique()
locations <- locations$Colnames

subset_full_2 <- transform(subset_full_1, Name = factor(Name, levels = legend)) %>%
        mutate(Name = replace_na(Name, "Other")) %>%
        unite("Selection", c("Location","Time"), sep = " - ", remove = FALSE) %>%
        transform(Selection = factor(Selection, levels = locations)) %>%
        filter(is.na(Selection)) %>%
        select(-Selection) %>%
        mutate(Time = str_replace(Time, pattern = "August", "Aug")) 

subset_all <- rbind(subset_downsampling, subset_full_2) %>%
        transform( Time = factor(Time, levels = c("April","June","Aug"))) %>%
        transform( Name = factor(Name, levels = legend))
          
      ggplot(data = subset_all, mapping = aes(x = Time, y = Percentage, fill = forcats::fct_rev(Name)))+
        geom_bar(stat="identity", position = "fill", width = 0.4)+
        facet_grid(cols = vars(Location))+ 
        labs(title = str_c("Most abundant ",ranks_plural), y = "Abundance")+
        theme_bw()+
        theme(plot.title = element_text(hjust = 0.5, size = 22))+
        theme(axis.text.x = element_text(size = 16))+
        theme(axis.text.y = element_text(size = 14))+
        theme(legend.text = element_text(size = 16))+
        theme(legend.title = element_blank())+
        scale_fill_manual(values = rev(brewer.pal(11,"Set3")))
      # save plot
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/family/Set1/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-adjusted_Main_fig_5_family_",classifyr,"_samples_V4.pdf"), width = 60, height = 15, units = "cm")
      ggsave(str_c(outfolder,"Relatve_abundance/Downsampled-dataset/family/Set1/Top10-highest-total-abundance/",date,"_Relative-abundance-plot_Individual-Most-abundant_level-adjusted_Main_fig_5_family_",classifyr,"_samples_V4.png"), width = 60, height = 15, units = "cm")
```

### Clustering & visualisation as heatmap 
```{r}

# create output directory
system(str_c("mkdir ",outfolder,"Heatmap"))
system(str_c("mkdir ",outfolder,"Heatmap/Downsampled-dataset/"))

# loop over all 4 downsampling sets
for (set in 1:4){
  
  # create specific output directories
  system(str_c("mkdir ",outfolder,"Heatmap/Downsampled-dataset/Set",set))

  # generate plots for downsampled data on genus and family level
  for (rank in c("family","genus")) {
    
    # Where we at?
    print(paste0('Processing downsampled data set ',set,' - on ',rank,' level'))
    
    # define input data depending in rank level
    if (rank == "family"){
      data.tmp <- filter(clean_classified_reads_downsampled[[1]][[set]], TaxRank == "F", Reads != "NA")%>%
      unite("Colnames", c("Location","Time"), sep = " - ") %>%
      select(Name, Colnames, Reads) %>%
      spread(key = Colnames, value = Reads)
    }
      
    if (rank == "genus"){
      data.tmp <- filter(clean_classified_reads_downsampled[[2]][[set]], TaxRank == "G", Reads != "NA")%>%
      unite("Colnames", c("Location","Time"), sep = " - ") %>%
      select(Name, Colnames, Reads) %>%
      spread(key = Colnames, value = Reads)
    } 
    
    # initialise data matrix for heatmap
    heatmap <- as.matrix(as.data.frame(data.tmp))
    rownames(heatmap) <- heatmap[,1]
    heatmap <- heatmap[,-1]
    class(heatmap) <- "numeric"
    
    # NA entries -> 0 entries
    heatmap[is.na(heatmap)] <- 0
    
    # remove rows (i.e. taxa) in which all entries -> 0
    heatmap <- heatmap[!apply(heatmap, 1, function(x) {all(x == 0)}),]
    
    # permform Log10(x+1) data transformation
    heatmap <- log10(heatmap+1)
    
    # make plots for all samples as well as samples+controls
    for (setting in c("samples","samples+controls")){
      if (rank == "family"){
        if (setting == "samples"){
          heatmap_1 <- heatmap[,1:(ncol(heatmap)-4)]
        } else {
          heatmap_1 <- heatmap[,1:(ncol(heatmap)-1)]
        }
      }   
      if (rank == "genus"){
        if (setting == "samples" ){
            heatmap_1 <- heatmap[,1:(ncol(heatmap)-3)]
        } else {
          heatmap_1 <- heatmap[,1:(ncol(heatmap))]
        }
      }   
    }
        
    # prepare annotations for histogram
    gAnnotationData <- colnames(heatmap_1)
    gAnnotationData <- tibble(gAnnotationData, .name_repair = ~ c("Name"))
    gAnnotationData$Location <- gAnnotationData$Name
    gAnnotationData <- gAnnotationData %>%
      separate(Location, into = c("Location","Time"), sep="-")
    gAnnotationData$Location <- trimws(gAnnotationData$Location)
    gAnnotationData$Time<- trimws(gAnnotationData$Time)
    
    gAnnotationData1 <- select(gAnnotationData, Location)%>%
      mutate(Location = str_replace(Location, pattern = "9.1", "9")) %>%
      mutate(Location = str_replace(Location, pattern = "9.2", "9"))
    gAnnotationData1$Location <- as_factor(gAnnotationData1$Location)
    gAnnotationData1<-as.data.frame(gAnnotationData1)
    row.names(gAnnotationData1) <- colnames(heatmap_1)
    
    gAnnotationData2 <- select(gAnnotationData, Location, Time)%>%
      mutate(Location = str_replace(Location, pattern = "9.1", "9")) %>%
      mutate(Location = str_replace(Location, pattern = "9.2", "9"))
    gAnnotationData1$Location <- as_factor(gAnnotationData1$Location)
    gAnnotationData2$Time <- as_factor(gAnnotationData2$Time)
    gAnnotationData2<-as.data.frame(gAnnotationData2)
    row.names(gAnnotationData2) <- colnames(heatmap_1)
    
    # define colors
    my_colour = list(
      Location = c("1" = "#2157A4", "2" = "#3694D1", "3" = "#65C6E8", "4" = "#9DD7ED", "5" = "#D6DE4F", "6" = "#FDCB44", "7" = "#F1861E", "8" = "#E63A11", "9" = "#a50026", "Mock" = "#636363", "N" = "#f0f0f0"),
      Time = c("April" = "#66c2a5", "June" = "#fc8d62", "August" = "#8da0cb"),
      Cluster = c("C1" = "#bdbdbd", "C2" = "#969696", "C3" = "#737373", "C4" = "#525252"))
    
     # perform hirachical clustering and devide taxa into 4 cluster
    my_hclust_gene <- hclust(dist(heatmap_1), method = "complete")
    as.dendrogram(my_hclust_gene) %>%
      plot(horiz = TRUE)
    my_gene_col <- cutree(tree = as.dendrogram(my_hclust_gene), k = 4)
    my_gene_col <- data.frame(Cluster = ifelse(test = my_gene_col == 1, yes = "C1", no = ifelse(test = my_gene_col == 2, yes = "C2", no= ifelse(test = my_gene_col == 3, yes = "C3", no = "C4"))))
    
    # plot heatmap
    my_heatmap <- pheatmap(heatmap_1, 
      clustering_distance_cols = "correlation",
      #clustering_distance_rows = "correlation",
      color = colorRampPalette(brewer.pal(n = 7, name = "Greens"))(100),
      show_rownames = FALSE, show_colnames = FALSE,
      annotation_row = my_gene_col,
      annotation_col = gAnnotationData2,
      annotation_colors = my_colour,
      cutree_rows = 4,
      cutree_cols = 2)
    
    # define function to save heatmap in .png format
    save_pheatmap_png <- function(x, filename, width=2000, height=2000, res = 300) {
      png(filename, width = width, height = height, res = res)
      grid::grid.newpage()
      grid::grid.draw(x$gtable)
      dev.off()
    }
    
    # define function to save heatmap in .pdf format
    save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
      stopifnot(!missing(x))
      stopifnot(!missing(filename))
      pdf(filename, width=width, height=height)
      grid::grid.newpage()
      grid::grid.draw(x$gtable)
      dev.off()
    }
    
    # save heatmap in both .png & .pdf file format
    save_pheatmap_png(my_heatmap, str_c(outfolder,"Heatmap/Downsampled-dataset/Set",set,"/",date,"_heatmap_downsampled_set-",set,"_",rank,"-level_",setting,".png"))
    save_pheatmap_pdf(my_heatmap, str_c(outfolder,"Heatmap/Downsampled-dataset/Set",set,"/",date,"_heatmap_downsampled_set-",set,"_",rank,"-level_",setting,".pdf"))
      
    # save cluster mapping
    cluster.mapping <- rownames_to_column(my_gene_col, var = "Name")
    save(cluster.mapping ,file = file.path(str_c(outfolder,"Heatmap/Downsampled-dataset/Set",set,"/",date,"_cluster-mapping_downsampled_set-",set,"_",rank,"-level_",setting,".RData")))
  }
}    
 
```

### Core microbiome
```{r}
# create output directory
system(str_c("mkdir ",outfolder,"Core_microbiome"))

# loop over family and genus downsampling datasets
for (rank in c("family","genus")) {
  if (rank == "family"){
    i <- 1
    Rank <- "F"
    threshold <- 0.1
  } else {
    i <- 2
    Rank <- "G"
    threshold <- 0.1
  }

  # loop over all 4 downsampling datasets
  for (set in 1:4){
    
    # Where we at?
    print(paste0('Processing downsampled data set ',set,' on ', rank,' level.'))
    
    # Prepare data for annotation 
    selection <- filter(clean_classified_reads_downsampled[[i]][[set]], TaxRank == Rank, Percentage != "NA", Location != "N" & Location != "Mock" & Location != "P") %>%
      mutate(Set = set)
    
    # combine data from all 4 sets  
    if (set == 1){
      core.microbiome <- selection
    } else {
      core.microbiome <- rbind(core.microbiome, selection)
    }
  }
  
  # calculate median and filter accordingly
  core.microbiome <- core.microbiome %>%
    group_by(Name) %>%
    mutate(Median = median(Percentage)) %>%
    ungroup(Name) %>%
    arrange(desc(Median)) %>%
    filter(Median >= threshold)
  
  # add clutster information from histogram of set 1
  load(file.path(str_c(outfolder,"Heatmap/Downsampled-dataset/Set1/",date,"_cluster-mapping_downsampled_set-1_",rank,"-level_samples+controls.RData")))
  core.microbiome <- left_join(core.microbiome, cluster.mapping)
  
  # plot either all or only core clusters
  for (setting in c("all.clusters","core.clusters")){
    if (setting == "core.clusters"){
      if (rank == "family"){
        core.microbiome <- filter(core.microbiome, Cluster != "C1" & Cluster != "C3") 
      }
      if (rank == "genus"){
        core.microbiome <- filter(core.microbiome, Cluster != "C1" & Cluster != "C4") 
      }
    }
    
    # correct levels 
    core.microbiome <- transform(core.microbiome, Cluster = factor(Cluster, levels = c("C4","C3","C2","C1")), Time = factor(Time, levels = c("April","June","August"))) 
    
    # make boxplot
    ggplot(core.microbiome,aes(x=reorder(Name, Percentage, FUN = median),y=Percentage)) +
      geom_boxplot(color = "black",fill="grey",show.legend = FALSE) +
      facet_grid(cols = vars(Time), rows = vars(Cluster), scales = "free_y",space = "free_y")+ #cols = vars(Time),
      scale_y_continuous(name ="Percentage", trans='log10', breaks = trans_breaks("log10", function(x) 10^x),labels = trans_format("log10", math_format(10^.x)))+
      coord_flip()+
      theme_minimal()+
      theme(axis.ticks = element_blank())+
      xlab("")
    
    # save plots
    ggsave(str_c(outfolder,"Core_microbiome/",date,"_Core-microbiome_V1_",rank,"-level_",classifyr,"_",setting,".pdf"), width = 20, height = 20, units = "cm")
    ggsave(str_c(outfolder,"Core_microbiome/",date,"_Core-microbiome_V1_",rank,"-level_",classifyr,"_",setting,".png"), width = 20, height = 20, units = "cm")
  
    # make violin plot  
    ggplot(core.microbiome,aes(x=reorder(Name, Percentage, FUN = median),y=Percentage)) +
      geom_violin(trim=FALSE, fill="gray")+
      facet_grid(cols = vars(Time),rows = vars(Cluster), scales = "free_y",space = "free_y")+ 
      scale_y_continuous(limits = c(0.01,100),name ="Percentage", trans='log10', breaks = trans_breaks("log10", function(x) 10^x),labels = trans_format("log10", math_format(10^.x)))+
      geom_hline(yintercept=threshold, linetype="dashed", color = "black")+
      coord_flip()+
      theme_bw()+
      theme(panel.grid.minor = element_blank())+
      xlab("")
  
    # save plots
    ggsave(str_c(outfolder,"Core_microbiome/",date,"_Core-microbiome_V2_",rank,"-level_",classifyr,"_",setting,".pdf"), width = 24, height = 25, units = "cm")
    ggsave(str_c(outfolder,"Core_microbiome/",date,"_Core-microbiome_V2_",rank,"-level_",classifyr,"_",setting,".png"), width = 24, height = 25, units = "cm")
  }
}
```

# Part 3: Differential taxa analysis 
## PCA analysis on the downsampled date
```{python, eval=FALSE}
use_condaenv("PuntSeq-analysis-env")

import os
import pandas as pd 
import numpy as np
import datetime
import matplotlib.pyplot as plt
from matplotlib.patches import Ellipse
import matplotlib.transforms as transforms
from sklearn.preprocessing import StandardScaler
from sklearn.decomposition import PCA

DATE = datetime.datetime.now().strftime('%Y-%m-%d')
outfolder = DATE + '_PuntSeq_Metagenomics-analysis/'
os.mkdir(outfolder+'PCA/')
# create folder for metagenomeSeq follow-up
os.mkdir(outfolder+'metagenomeSeq/')

def confidence_ellipse(x, y, ax, n_std=2.0, facecolor='none', alpha=0.5, **kwargs):
    """
    Create a plot of the covariance confidence ellipse of `x` and `y`

    Parameters
    ----------
    x, y : array_like, shape (n, )
        Input data.

    ax : matplotlib.axes.Axes
        The axes object to draw the ellipse into.

    n_std : float
        The number of standard deviations to determine the ellipse's radiuses.

    Returns
    -------
    matplotlib.patches.Ellipse

    Other parameters
    ----------------
    kwargs : `~matplotlib.patches.Patch` properties
    """
    if x.size != y.size:
        raise ValueError("x and y must be the same size")

    cov = np.cov(x, y)
    pearson = cov[0, 1]/np.sqrt(cov[0, 0] * cov[1, 1])
    # Using a special case to obtain the eigenvalues of this
    # two-dimensionl dataset.
    ell_radius_x = np.sqrt(1 + pearson)
    ell_radius_y = np.sqrt(1 - pearson)
    ellipse = Ellipse((0, 0),
        width=ell_radius_x * 2,
        height=ell_radius_y * 2,
        facecolor=facecolor, alpha=alpha, linestyle='None', edgecolor=facecolor,
        **kwargs)

    # Calculating the stdandard deviation of x from
    # the squareroot of the variance and multiplying
    # with the given number of standard deviations.
    scale_x = np.sqrt(cov[0, 0]) * n_std
    mean_x = np.mean(x)

    # calculating the stdandard deviation of y ...
    scale_y = np.sqrt(cov[1, 1]) * n_std
    mean_y = np.mean(y)

    transf = transforms.Affine2D() \
        .rotate_deg(45) \
        .scale(scale_x, scale_y) \
        .translate(mean_x, mean_y)

    ellipse.set_transform(transf + ax.transData)
    return ax.add_patch(ellipse)
    

def autolabel_feat(rects, R_df, ax):
    """
    Attach a text label above each bar displaying its height
    """
    m = 0
    for rect in rects:
        width = rect.get_width()
        if width != 0:
            ax.text(0.01, rect.get_y() + rect.get_height()/10. ,'%s' % R_df.index[m], ha='left', va='bottom')

        m += 1

def autolabel2_feat(rects, R_df, ax):
    """
    Attach a text label above each bar displaying its height
    """
    m = 0
    for rect in rects:
        width = rect.get_width()
        if width != 0:
            ax.text(-0.01, rect.get_y() + rect.get_height()/10. ,'%s' % R_df.index[m], ha='right', va='bottom')

        m += 1
        
        

### load one downsampled dataset and prepare PCA
nr = 1 
data = pd.read_csv(outfolder+'Downsampling/genus/final_%s.tsv' %nr, sep='\t', index_col=0, header=0)

# plot PCA of all remaining samples (>30k reads)
x = data.copy()

# save positive controls in different dataframes to plot them on top of the PCA later
x2 = x[['Mock-April_genus.tsv','Mock-June_genus.tsv','Mock-August_genus.tsv']]

# delte positive and negative controls
del x['N-August_genus.tsv']
del x['Mock-April_genus.tsv']; del x['Mock-August_genus.tsv']; del x['Mock-June_genus.tsv']

datacolumns = x.columns
dataindex = x.index

# transpose to have the features in the column of the dataframe
x = x.transpose()
x2 = x2.transpose()

# logarithmise and standardise the features by removing the mean and scaling to unit variance
x = np.log10(x + 1)
x = StandardScaler().fit_transform(x)

x2 = np.log10(x2 + 1)
x2 = StandardScaler().fit_transform(x2)

# save matrix for follow-up metagenomeSeq analysis
pd.DataFrame(x, index = datacolumns, columns = dataindex).transpose().to_csv('metagenomeSeq/fitfeature.tsv', sep='\t')

# define a PCA with 10 components and 
pca = PCA(n_components=10)

# fit the PCA based on all water samples ('x'), the only transform the positive controls ('x2')
principalComponents = pca.fit_transform(x)
principalComponents2 = pca.transform(x2)
principalDf = pd.DataFrame(data = principalComponents, columns = ['PC 1', 'PC 2', 'PC 3', 'PC 4', 'PC 5', 'PC 6', 'PC 7', 'PC 8', 'PC 9', 'PC 10'])

# define target (location) and time point (month) per sample
principalDf['target'] = datacolumns
principalDf['month'] = [x.split('-')[1].split('_')[0] for x in principalDf.target.values]

# transform data
principalDf.index = principalDf.month
principalDf['targetnice'] = [x.split('-')[0] for x in principalDf.target.values]
b, c = principalDf.iloc[1], principalDf.iloc[2]
temp = principalDf.iloc[1].copy()
principalDf.iloc[1] = c
principalDf.iloc[2] = temp

principalComponents2 = pd.DataFrame(data = principalComponents2, columns = ['PC 1', 'PC 2', 'PC 3', 'PC 4', 'PC 5', 'PC 6', 'PC 7', 'PC 8', 'PC 9', 'PC 10'])
principalComponents2['target'] = ['Mock-April_genus.tsv','Mock-June_genus.tsv','Mock-August_genus.tsv']
principalComponents2['month'] = ['April','June','August']
principalComponents2['targetnice'] = ['P','P','P']
principalComponents2.index = ['April','June','August']
principalDf = pd.concat([principalDf, principalComponents2])


### plot PCA 

plt.style.use('classic')

# choose PCs
pc1 = 1
pc2 = 2

# plot and add percentage variance explained to the axes
n1 = 'PC %i' %pc1
n2 = 'PC %i' %pc2
var1 = pca.explained_variance_ratio_[pc1-1]
var2 = pca.explained_variance_ratio_[pc2-1]

fig, ax = plt.subplots(figsize=(8, 6), facecolor='white')
ax.set_xlabel('%s (Variance = %s)' % (n1,'%.3f'%(var1)), fontsize = 12)
ax.set_ylabel('%s (Variance = %s)'% (n2, '%.3f'%(var2)), fontsize = 12)

for target, color in zip(['April','June','August'], ['#66c2a5','#fc8d62', '#8da0cb']):
    indicesToKeep = principalDf['month'] == target
    ellip = np.array(principalDf.loc[indicesToKeep, [n1,n2]])
    ell = confidence_ellipse(ellip[:,0], ellip[:,1], ax=ax, facecolor=color, alpha=0.3)
    plt.scatter(np.mean(ellip[:,0]), np.mean(ellip[:,1]), facecolor=color, s=10)
    ax.scatter(principalDf.loc[indicesToKeep, n1]
               , principalDf.loc[indicesToKeep, n2]
               , c = color
               , s = 50
               , edgecolors='none')
    
for label, x, y in zip(principalDf.targetnice.values, principalDf[n1].values, principalDf[n2].values):
    plt.annotate(
        label,
        xy=(x, y), xytext=(-6, -8),
        textcoords='offset points', ha='right', va='bottom',
        fontsize=10)  

ax.legend(('April', 'June', 'August'), scatterpoints=1, markerscale=1.2,borderpad=0.5,labelspacing=1,loc=1,fontsize=12, ncol=1,fancybox=True,shadow=False,frameon=False,numpoints=1)
plt.axhline(0, c='black')
plt.axvline(0, c='black')
ax.set_axis_bgcolor('white')
ax.spines['right'].set_visible(False)
ax.spines['top'].set_visible(False)
ax.yaxis.set_ticks_position('left')
ax.xaxis.set_ticks_position('bottom')


fig.savefig(outfolder+'PCA/PCA_%i_%i.pdf' %(pc1,pc2), bbox_inches='tight')


### 
scipy.stats.kruskal(df['PC 3'].loc['April'],df['PC 3'].loc['August'],
  df['PC 3'].loc['June'])
scipy.stats.mannwhitneyu(df['PC 3'].loc['June'],df['PC 3'].loc['April'],
  alternative = 'two-sided')
  scipy.stats.mannwhitneyu(df['PC 3'].loc['June'],df['PC 3'].loc['August'],
  alternative = 'two-sided')
  scipy.stats.mannwhitneyu(df['PC 3'].loc['August'],df['PC 3'].loc['April'],
  alternative = 'two-sided')


### examine the contributions of bacteria to the PCs
# choose PCs
pc1 = 1
pc2 = 2
n1 = 'PC %i' %pc1
n2 = 'PC %i' %pc2

datap = pd.DataFrame(pca.components_[[pc1-1,pc2-1],], index = [n1,n2], columns = data.index)
datap = datap.transpose()

# sort according to one PC
pc_c = n1
dataps = datap.copy()  
dataps['PCabs'] = dataps[pc_c].abs()         
dataps = dataps.sort('PCabs', ascending=False)    
# look at top 10 contributing bacteria
dataps[pc_c].head(n=10)      

# plot contributions as a barplot
n_bacteria = 10
dataplot = dataps[pc_c].head(n=n_bacteria)   

negative_data = []
positive_data = []
negative_data_std = []
positive_data_std = []
for feature in dataplot.index:
    if dataplot.loc[feature] >= 0:
        positive_data.append(dataplot.loc[feature])
        negative_data.append(0)
    elif dataplot.loc[feature] < 0:
        negative_data.append(dataplot.loc[feature])
        positive_data.append(0)
        
x = range(n_bacteria)[::-1]
fig = plt.figure(figsize=(6, 10), facecolor='white')
ax = plt.subplot(111)
rects1= ax.barh(x, negative_data, align='edge', color='sienna', ecolor='black', alpha=0.7, capsize=4) 
rects2 = ax.barh(x, positive_data, align='edge', color='cornflowerblue', ecolor='black',alpha=0.7, capsize=4) 
ax.spines['right'].set_visible(False)
ax.spines['top'].set_visible(False)
ax.spines['left'].set_visible(False)
ax.yaxis.set_visible(False)
plt.axvline(x=0., color="black")
ax.yaxis.set_ticks_position('left')
ax.xaxis.set_ticks_position('bottom')


autolabel_feat(rects1, dataplot[:n], ax)
autolabel2_feat(rects2, dataplot[:n], ax)
    
ax.set_xlabel('contribution to %s' %pc_c)
fig.savefig(outfolder+'PCA/PCA_contributions_%s.pdf' %pc_c, bbox_inches='tight')





### apply PCA to ionics data (also available on github)
url = "https://raw.githubusercontent.com/d-j-k/puntseq/master/Ionics_Table.xlsx"
s = requests.get(url).content
data = pd.read_excel(io.StringIO(s.decode('utf-8')), 'Environmental metadata (sample)')

# clean up dataframe
data.columns = data.iloc[1,:]
data.columns.values[5:] = data.iloc[2,5:]
data = data.drop([0,1,2])
data['month'] = 'April'
data['month'][data.Batch==2] = 'June'
data['month'][data.Batch==3] = 'August'
data['targetnice'] = data.Barcode
targetnice = data.targetnice.values
month = data.month.values
# only save ionics in actual dataframe, delete all metadata
data = data.iloc[:,4:]
x = data.iloc[:,:-2]

# delete Fe, Mn and Al (can't be trusted)
del x['Fe']; del x['Mn']; del x['Al']

datacolumns = x.columns 

# logarithmise and standardise dataframe
x = x.astype(np.float64)
x = np.log10(x + 1)
x = StandardScaler().fit_transform(x)

# initiate PCA with 10 PCs
pca = PCA(n_components=10)
principalComponents = pca.fit_transform(x)
principalDf = pd.DataFrame(data = principalComponents, columns = ['PC 1', 'PC 2', 'PC 3', 'PC 4', 'PC 5', 'PC 6', 'PC 7', 'PC 8', 'PC 9', 'PC 10'])
principalDf['targetnice'] = targetnice
principalDf['month'] = month
principalDf.index = principalDf.month

# choose PCs
pc1 = 1
pc2 = 2
n1 = 'PC 1'
n2 = 'PC 2'
var1 = pca.explained_variance_ratio_[pc1-1]
var2 = pca.explained_variance_ratio_[pc2-1]

fig, ax = plt.subplots(figsize=(8, 6), facecolor='white')
ax.set_xlabel('%s (Variance = %s)' % (n1,'%.3f'%(var1)), fontsize = 12)
ax.set_ylabel('%s (Variance = %s)'% (n2, '%.3f'%(var2)), fontsize = 12)

for target, color in zip(['April','June','August'], ['#66c2a5','#fc8d62', '#8da0cb']):
    indicesToKeep = principalDf['month'] == target
    ellip = np.array(principalDf.loc[indicesToKeep, [n1,n2]])
    confidence_ellipse(ellip[:,0], ellip[:,1], ax=ax, facecolor=color, alpha=0.3)
    ax.scatter(principalDf.loc[indicesToKeep, n1]
               , principalDf.loc[indicesToKeep, n2]
               , c = color
               , s = 50
               , edgecolors='none')

for label, x, y in zip(principalDf.targetnice.values, principalDf[n1].values, principalDf[n2].values):
    plt.annotate(
        label,
        xy=(x, y), xytext=(-6, -8),
        textcoords='offset points', ha='right', va='bottom',
        fontsize=10)  
    
ax.legend(('April', 'June', 'August'), scatterpoints=1, markerscale=1.2,borderpad=0.5,labelspacing=1,loc=2,fontsize=12, ncol=1,fancybox=True,shadow=False,frameon=False,numpoints=1)
plt.axhline(0, c='black')
plt.axvline(0, c='black')
ax.set_axis_bgcolor('white')
ax.spines['right'].set_visible(False)
ax.spines['top'].set_visible(False)
ax.yaxis.set_ticks_position('left')
ax.xaxis.set_ticks_position('bottom')

fig.savefig(outfolder+'PCA/PCA_ionics_%i_%i.pdf' %(pc1,pc2), bbox_inches='tight')



### asssess contributions to PCs 
# choose PCs
pc1 = 1
pc2 = 2
n1 = 'PC 1'
n2 = 'PC 2'
datap = pd.DataFrame(pca.components_[[pc1-1,pc2-1],], index = [n1,n1], columns = datacolumns)
datap = datap.transpose()
n = datap.shape[0]       

# show contributions of all ionics to chosen PC
pc_c = n1
dataps = datap.copy()  
dataps['PCabs'] = dataps[pc_c].abs()         
dataps = dataps.sort('PCabs', ascending=False)    
dataps[pc_c].head(n=n)      

dataion = dataps[pc_c].head(n=n)     

# transform features and plot barplot of contributions
negative_data = []
positive_data = []
negative_data_std = []
positive_data_std = []
for feature in dataion.index:
    if dataion.loc[feature] >= 0:
        positive_data.append(dataion.loc[feature])
        negative_data.append(0)
    elif dataion.loc[feature] < 0:
        negative_data.append(dataion.loc[feature])
        positive_data.append(0)

x = range(n)[::-1]

fig = plt.figure(figsize=(6, 10), facecolor='white')
ax = plt.subplot(111)
rects1 = ax.barh(x, negative_data, align='edge', color='sienna', ecolor='black', alpha=0.7, capsize=4) 
rects2 = ax.barh(x, positive_data, align='edge', color='cornflowerblue', ecolor='black',alpha=0.7, capsize=4) 
ax.spines['right'].set_visible(False)
ax.spines['top'].set_visible(False)
ax.spines['left'].set_visible(False)
ax.yaxis.set_visible(False)
plt.axvline(x=0., color="black")
ax.yaxis.set_ticks_position('left')
ax.xaxis.set_ticks_position('bottom')

autolabel_feat(rects1, dataion[:n], ax)
autolabel2_feat(rects2, dataion[:n], ax)
    
ax.set_xlabel('contribution to %s' %pc_c)
fig.savefig(outfolder+'PCA/PCA_ionics_contributions_%s.pdf' %pc_c, bbox_inches='tight')

```

## Apply metagenomeSeq for seasonal divergence analysis
```{r eval=FALSE}
# install metagenome Seq package
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("metagenomeSeq")

# load library
library(metagenomeSeq)

# load data that was created during PCA analysis
x = read.table('metagenomeSeq/fitfeature.tsv', sep='\t', header = TRUE, row.names = 1)
ph = c('April', 'August', 'June', 'April',
       'August', 'April', 'April', 'August', 'April',
       'August', 'August', 'June', 'April', 'April',
       'August', 'August')
phenoData=AnnotatedDataFrame(data.frame(ph, row.names = colnames(x)))

# Group June and August samples to compare spring versus summer months:
# rename June to August samples
ph = c('April', 'August', 'August', 'April',
       'August', 'April', 'April', 'August', 'April',
       'August', 'August', 'August', 'April', 'April',
       'August', 'August')
phenoData=AnnotatedDataFrame(data.frame(ph, row.names = colnames(x)))

# define new metagenomeSeq experiment
obj = newMRexperiment(as.matrix(x), phenoData = phenoData) 

pd <- pData(obj)

# define model
mod <- model.matrix(~1 + ph, data = pd)

# calculates the percentile for which to sum counts up to and scale by
p = cumNormStatFast(obj, rel=0.1)
obj = cumNorm(obj, p = p)

# fitfeature model
res1 = fitFeatureModel(obj, mod)
head(MRcoefs(res1))

# assess nominally significant bacterial families
MRcoefs(res1)[rownames(MRcoefs(res1)[MRcoefs(res1)$pvalues < 0.05,]),]

```

## Potentially waterborne pathogens and genera associated with wasterwater
```{r}
# create output directory
system(str_c("mkdir ",outfolder,"Interesting_taxa"))

# (down-)load lists of taxa to investigate

# EU+WHO drinking water pathogens indicators
infile_taxa_EUWHO <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/interesting_taxa/2019-11-25_EU%2BWHO_drinking_water_pathogens%2Bindicators.txt")
destfile_taxa_EUWHO=str_c(outfolder,"Interesting_taxa/2019-11-25_EU+WHO_drinking_water_pathogens+indicators.txt")
if(!file.exists(destfile_taxa_EUWHO)){
  res <- tryCatch(download.file(infile_taxa_EUWHO, destfile_taxa_EUWHO, method="auto"), error=function(e) 1)
}

# potential waterborne pathogens (Jin et al 2018 Scientific Reports)
infile_taxa_Jin <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/interesting_taxa/2019-11-26_Jin2018_ScientificReports_potential_waterborne_pathogens.txt")
destfile_taxa_Jin=str_c(outfolder,"Interesting_taxa/2019-11-26_Jin2018_ScientificReports_potential_waterborne_pathogens.txt")
if(!file.exists(destfile_taxa_Jin)){
  res <- tryCatch(download.file(infile_taxa_Jin, destfile_taxa_Jin, method="auto"), error=function(e) 1)
}

# human pathogens (PATRIC database manually curated)
infile_taxa_PATRIC <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/interesting_taxa/2019-11-25_PATRIC_database_manually_curated_pathogens.txt")
destfile_taxa_PATRIC=str_c(outfolder,"Interesting_taxa/2019-11-25_PATRIC_database_manually_curated_pathogens.txt")
if(!file.exists(destfile_taxa_PATRIC)){
  res <- tryCatch(download.file(infile_taxa_PATRIC, destfile_taxa_PATRIC, method="auto"), error=function(e) 1)
}

# wastewater core taxa (Wu et al 2019 Nature Microbiology)
infile_taxa_Wu <- str_c("https://raw.githubusercontent.com/d-j-k/puntseq/master/downstream_analysis/interesting_taxa/2019-11-25_Wu2019_NatureMicrobiology_wastewater_coretaxa.txt")
destfile_taxa_Wu=str_c(outfolder,"Interesting_taxa/2019-11-25_Wu2019_NatureMicrobiology_wastewater_coretaxa.txt")
if(!file.exists(destfile_taxa_Wu)){
  res <- tryCatch(download.file(infile_taxa_Wu, destfile_taxa_Wu, method="auto"), error=function(e) 1)
}

# merge genus information from all lists abd adjust to SILVA annotation where nessesary
EUWHO_raw <- read_tsv(destfile_taxa_EUWHO)
EUWHO <- select(EUWHO_raw, -Family, -X7, -X8) %>%
  mutate(Genus = str_replace(Genus, pattern = "Burkholderia", "Burkholderia-Caballeronia-Paraburkholderia")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Escherichia", "Escherichia-Shigella")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Hafnia", "Hafnia-Obesumbacterium")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Mycobacteria", "	Mycobacterium")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Clostridium", "Clostridium sensu stricto 1")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Chlamydophila", "Chlamydia")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Borrelia", "Borreliella"))%>%
  filter(Genus != "Shigella") %>%
  rename(EU_BW = EU_bathing_water_indicators)%>%
  rename(WHO_DWP = WHO_drinking_water_pathogens)%>%
  rename(WHO_pDWP = WHO_pot_drinking_water_pathogens)
  
Jin_raw <- read_tsv(destfile_taxa_Jin)
Jin <- select(Jin_raw, Genus) %>%
  filter(Genus != "NA")%>%
  unique()%>%
  mutate( Jin = "Y")

PATRIC_raw <- read_tsv(destfile_taxa_PATRIC)
PATRIC <- select(PATRIC_raw, Genus)%>%
  unique()%>%
  mutate(PATRIC = "Y")%>%
  mutate(Genus = str_replace(Genus, pattern = "Burkholderia", "Burkholderia-Caballeronia-Paraburkholderia")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Escherichia", "Escherichia-Shigella")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Hafnia", "Hafnia-Obesumbacterium")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Mycobacteria", "	Mycobacterium")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Clostridium", "Clostridium sensu stricto 1")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Chlamydophila", "Chlamydia")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Borrelia", "Borreliella"))%>%
  filter(Genus != "Shigella")

Wu_raw <- read_tsv(destfile_taxa_Wu)
Wu <- select(Wu_raw, Genus)%>%
  filter(Genus != "NA") %>%
  mutate(Wu = "Y") %>%
  mutate(Genus = str_replace(Genus, pattern = "Burkholderia", "Burkholderia-Caballeronia-Paraburkholderia")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Escherichia", "Escherichia-Shigella")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Hafnia", "Hafnia-Obesumbacterium")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Mycobacteria", "	Mycobacterium")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Clostridium", "Clostridium sensu stricto 1")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Chlamydophila", "Chlamydia")) %>%
  mutate(Genus = str_replace(Genus, pattern = "Borrelia", "Borreliella"))%>%
  filter(Genus != "Shigella")

merge1 <- full_join(EUWHO, Jin)
merge2 <- full_join(merge1, PATRIC)
interesting.taxa <- full_join(merge2, Wu)

# adjust taxonomie to silva nomenclature 
interesting.taxa <- rename(interesting.taxa, Name = Genus)
  

# convert NA values
interesting.taxa[is.na(interesting.taxa)] <- "N"
interesting.taxa <- mutate(interesting.taxa, Wastewater_only = ifelse(EU_BW == "N" & coliforms == "N" & WHO_DWP =="N" & WHO_pDWP =="N" & Jin == "N" & PATRIC == "N" & Wu =="Y", "Y", "N")) 

# map with read data
interesting.taxa.reads <- left_join(interesting.taxa, clean_classified_reads_downsampled[[2]][[1]])
interesting.taxa.reads <- interesting.taxa.reads %>%
  filter(Reads != "NA")
  
# convert data into matrix format
interesting.taxa.reads.clean <- unite(interesting.taxa.reads, "Colnames", c("Location","Time"), sep = " - ") %>%
  select(Name, Colnames, Reads) %>%
  spread(key = Colnames, value = Reads)

# add max count and set all genera with max count < 10 to 0
list_selection <- interesting.taxa.reads.clean
list_selection$max <- apply(interesting.taxa.reads.clean[, 2:16], 1, max)
for (i in c(1:nrow(list_selection))){
  if(list_selection[i,20] < 10) {
    for (j in c(2:16)){
      list_selection[i,j] = 0 
    }
  }  
}

# normalise mean of 0 and sd of 1
list_toscale <- list_selection[1:16]
scaled_rows = t( scale(t(list_toscale[2:16] )))
  
# check that we get mean of 0 and sd of 1
rowMeans(scaled_rows)  # faster version of apply(scaled.dat, 2, mean)
apply(scaled_rows, 1, sd)

# combine normalised (scaled) list with genus information again  
list_scaled<- bind_cols(list_toscale[1], data.frame(scaled_rows))
  
# convert into plottable form
list_normalised <- gather(list_scaled, key = "Location", value = "Reads_normalised", 2:16) %>%
  mutate(Location = str_replace(Location, pattern = "X", "")) %>%
  mutate(Location = str_replace(Location, pattern = "\\.\\.\\.", " - "))
  
# extract read data and calculate log10(read+1)
interesting.taxa.reads.tidy <- gather(interesting.taxa.reads.clean, key = "Location", value = "Reads", 2:19) %>% 
  filter(Location != "Mock - April" & Location != "Mock - June" & Location != "Mock - August") 
interesting.taxa.reads.tidy$logreadsp1 <- log2(interesting.taxa.reads.tidy$Reads+1)

# join lists  
interesting.taxa.final.list <- full_join(interesting.taxa.reads.tidy, list_normalised ) 
interesting.taxa.final.list <- left_join(interesting.taxa.final.list, interesting.taxa)
interesting.taxa.final.list <- filter(interesting.taxa.final.list, Reads_normalised != "NaN") %>%
  transform(Location = factor(Location, levels = c("1 - April","1 - June","1 - August","2 - April","2 - August","3 - April","4 - April","4 - August","5 - April","5 - August","6 - June","6 - August","7 - April","8 - April","9.1 - August")))
  
  
# make boxplot
ggplot(interesting.taxa.final.list, aes(reorder(Name, Reads, FUN = median), Reads+1)) +
  #geom_jitter(aes(colour = Location),show.legend = FALSE)+
  geom_boxplot(color = "black",fill="grey",show.legend = FALSE) +
  facet_grid(rows = vars(Wastewater_only), scales = "free_y",space = "free_y")+ 
  scale_y_continuous(name ="Reads",trans=log10_trans(), breaks=c(10, 100, 1000, 10000), label = comma)+
  coord_trans(y="log10")+
  geom_hline(yintercept=10, linetype="dashed", color = "black")+
  coord_flip()+
  theme_minimal()+
  theme(axis.ticks = element_blank())+
  xlab("")
  
  ggsave(str_c(outfolder,"Interesting_taxa/",date,"Pathogens_family-level_bubble-chart_1.pdf"), width = 12, height = 20, units = "cm")
  ggsave(str_c(outfolder,"Interesting_taxa/",date,"Pathogens_family-level_bubble-chart_1.png"), width = 12, height = 20, units = "cm")
  
  #normalised dispersion from mean SD
  ggplot(interesting.taxa.final.list, aes(x=factor(Location), y=reorder(Name, Reads, FUN = median), size=abs(log(Reads+1)), fill=Reads_normalised)) +
    geom_point(shape = 21,alpha=1, colour="grey") +    # plot as points
    facet_grid(rows = vars(Wastewater_only), scales = "free_y",space = "free_y")+ 
    scale_size(range = c(0.1, 7.5),name = "σ")+
    scale_fill_gradient2(name = "σ", trans="reverse")+
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust = 1),
          axis.text.y = element_blank(),
          axis.line = element_blank(),            # disable axis lines
          axis.ticks = element_blank(),
          axis.title = element_blank(),           # disable axis titles
          panel.border = element_blank(),         # disable panel border
          panel.grid.major.x = element_blank(),   # disable lines in grid on X-axis
          panel.grid.minor.x = element_blank())   # disable lines in grid on X-axis
  
  ggsave(str_c(outfolder,"Interesting_taxa/",date,"Pathogens_family-level_bubble-chart_2.pdf"), width = 18, height = 20, units = "cm")
  ggsave(str_c(outfolder,"Interesting_taxa/",date,"Pathogens_family-level_bubble-chart_2.png"), width = 18, height = 20, units = "cm")
  
# plot metadata WHO_DWP annotation
ggplot(interesting.taxa.final.list, aes(x=reorder(Name, Reads, FUN = median), y=0.5, fill=WHO_DWP))+
  geom_tile(color="white", size=0.1) +
  facet_grid(rows = vars(Wastewater_only), scales = "free_y",space = "free_y")+ 
  scale_fill_manual(values = c("Y"="#fbb4ae", "N"="gray90")) +
  coord_flip()+
  labs(x=NULL, y=NULL, title=NULL)+
  theme_minimal()+
  theme(axis.ticks = element_blank())+
  theme(panel.grid = element_blank())+
  theme(axis.text.x=element_blank())
ggsave(str_c(outfolder,"Interesting_taxa/",date,"WHO_DWP_map.pdf"), width = 9, height = 20, units = "cm")

# plot metadata WHO_pDWP annotation
ggplot(interesting.taxa.final.list, aes(x=reorder(Name, Reads, FUN = median), y=0.5, fill=WHO_pDWP))+
  geom_tile(color="white", size=0.1) +
  facet_grid(rows = vars(Wastewater_only), scales = "free_y",space = "free_y")+ 
  scale_fill_manual(values = c("Y"="#fbb4ae", "N"="gray90")) +
  coord_flip()+
  labs(x=NULL, y=NULL, title=NULL)+
  theme_minimal()+
  theme(axis.ticks = element_blank())+
  theme(panel.grid = element_blank())+
  theme(axis.text.x=element_blank())
ggsave(str_c(outfolder,"Interesting_taxa/",date,"WHO_pDWP_map.pdf"), width = 9, height = 20, units = "cm")

# plot metadata Jin annotation
ggplot(interesting.taxa.final.list, aes(x=reorder(Name, Reads, FUN = median), y=0.5, fill=Jin))+
  geom_tile(color="white", size=0.1) +
  facet_grid(rows = vars(Wastewater_only), scales = "free_y",space = "free_y")+ 
  scale_fill_manual(values = c("Y"="#b3cde3", "N"="gray90")) +
  coord_flip()+
  labs(x=NULL, y=NULL, title=NULL)+
  theme_minimal()+
  theme(axis.ticks = element_blank())+
  theme(panel.grid = element_blank())+
  theme(axis.text.x=element_blank())
ggsave(str_c(outfolder,"Interesting_taxa/",date,"Jin_map.pdf"), width = 9, height = 20, units = "cm")

# plot metadata Wu annotation
ggplot(interesting.taxa.final.list, aes(x=reorder(Name, Reads, FUN = median), y=0.5, fill=Wu))+
  geom_tile(color="white", size=0.1) +
  facet_grid(rows = vars(Wastewater_only), scales = "free_y",space = "free_y")+ 
  scale_fill_manual(values = c("Y"="#ccebc5", "N"="gray90")) +
  coord_flip()+
  labs(x=NULL, y=NULL, title=NULL)+
  theme_minimal()+
  theme(axis.ticks = element_blank())+
  theme(panel.grid = element_blank())+
  theme(axis.text.x=element_blank())
ggsave(str_c(outfolder,"Interesting_taxa/",date,"Wu_map.pdf"), width = 9, height = 20, units = "cm")

# plot metadata PATRIC annotation
ggplot(interesting.taxa.final.list, aes(x=reorder(Name, Reads, FUN = median), y=0.5, fill=PATRIC))+
  geom_tile(color="white", size=0.1) +
  facet_grid(rows = vars(Wastewater_only), scales = "free_y",space = "free_y")+ 
  scale_fill_manual(values = c("Y"="#decbe4", 'N' ="gray90")) +
  coord_flip()+
  labs(x=NULL, y=NULL, title=NULL)+
  theme_minimal()+
  theme(axis.ticks = element_blank())+
  theme(panel.grid = element_blank())+
  theme(axis.text.x=element_blank())
ggsave(str_c(outfolder,"Interesting_taxa/",date,"PATRIC_map.pdf"), width = 9, height = 20, units = "cm")


# relative abundance of reads mapping to potentially pathogeneic genera per location
pot.pathogens.rel.abundance <- select(interesting.taxa.final.list, Location, Reads, Wastewater_only) %>%
  filter(Wastewater_only == "N") %>%
  group_by(Location) %>%
  summarize(Sum = sum(Reads, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(rel_abundance = (Sum/37000)*100)

# plot relativ abundance mapping to genera harboring pot. pathogeneic bacteria
ggplot(pot.pathogens.rel.abundance, aes(x=Location, y=0.5, fill=rel_abundance))+
  geom_tile(color="white", size=0.1) +
  scale_fill_gradient(low = "#feedde", high = "#d94701")+
  labs(x=NULL, y=NULL, title=NULL)+
  geom_text(aes(label = paste0(round(rel_abundance,1))), position = position_stack(vjust = 0.5), color = "black")+
  theme_minimal()+
  theme(axis.ticks = element_blank())+
  theme(panel.grid = element_blank())+
  theme(axis.text.y=element_blank())
ggsave(str_c(outfolder,"Interesting_taxa/",date,"relativ_abundance_map.pdf"), width = 18, height = 2, units = "cm")


```


# Part 4: Additional analyses
## Hydrochemistry plots
```{r eval=FALSE}

# create output directory
system(str_c("mkdir ",outfolder,"Hydrochemistry"))


# 1. Gaillardet et al., Chemical Geology 1999
#############################################

# a. Import their table
gaillardet <- read.table(str_c(outfolder,"Supplementary_data/Gaillardet_1999_Table1.txt"), header = T,
                              sep = '\t', check.names = F, fill = T)
for(i in 1:ncol(gaillardet)){
  gaillardet[,i] <- gsub(' ', '', gaillardet[,i])
}

# b. Add Ionics measurements to it
puntseq <- read_xlsx(str_c(outfolder,"Supplementary_data/Supplementary_Table_S1.xlsx"), sheet = 2, skip = 3)[,-c(2,4:6)]
puntseq <- as.data.frame(puntseq)
puntseq[,1] <- gsub('1', 'April', puntseq[,1]); puntseq[,1] <- gsub('2', 'June', puntseq[,1]); puntseq[,1] <- gsub('3', 'August', puntseq[,1])
rownames(puntseq) <- paste(puntseq[,1], puntseq[,2], sep = ' - ')
puntseq <- puntseq[,-c(1:2)]

# c. Build ratios
gaillardet.HCO3_Na_vs_Ca_Na <- cbind(as.numeric(as.character(gaillardet[,"HCO3 (μmol l−1)"]))/as.numeric(as.character(gaillardet[,"Na (μmol l−1)"])),
                                     as.numeric(as.character(gaillardet[,"Ca (μmol l−1)"]))/as.numeric(as.character(gaillardet[,"Na (μmol l−1)"])),
                                     as.numeric(as.character(gaillardet[,"TDS (mg l−1)"])))
rownames(gaillardet.HCO3_Na_vs_Ca_Na) <- rownames(gaillardet)
colnames(gaillardet.HCO3_Na_vs_Ca_Na) <- c('HCO3/Na', 'Ca/Na', 'TPS')

puntseq.HCO3_Na_vs_Ca_Na <- cbind(as.numeric(as.character(puntseq[,"HCO3 (calculated)"]))/as.numeric(as.character(puntseq[,"Na"])),
                                  as.numeric(as.character(puntseq[,"Ca"]))/as.numeric(as.character(puntseq[,"Na"])))
colnames(puntseq.HCO3_Na_vs_Ca_Na) <- colnames(gaillardet.HCO3_Na_vs_Ca_Na)[1:2]
rownames(puntseq.HCO3_Na_vs_Ca_Na) <- rownames(puntseq)

# d. Log/Log plot (0.1 - 20 (x) / 50 (y))

## Fig. S3e
pdf(str_c(outfolder,"Hydrochemistry/",date,"_Carbonation_world_rivers_S3e.pdf"), width = 11, height = 7)
transp <- 0.7
par(mar = c(7,9,2,2))

## gaillardet main data
plot(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,1], 
     x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,2], log = 'xy', 
     xlim = c(0.1, 20), ylim = c(0.1, 50), pch = 21, 
     bg = 'white', col = 'black', ylab = "", xlab = "",
     cex.lab = 3.5, main = "", cex.main = 3, cex = 2, bty = 'n',
     cex.axis = 2)
mtext(text = "Ca/Na", side = 1, line = 5, cex = 3.5)
mtext(text = "HCO3/Na", side = 2, line = 5, cex = 3.5)

## gaillardet polluted rivers data
points(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,1], x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,2],
       cex = 2, pch = 21, bg = 'white', col = 'black')

## April ellipse
dataEllipse(x = puntseq.HCO3_Na_vs_Ca_Na[1:10,2],
            y = puntseq.HCO3_Na_vs_Ca_Na[1:10,1],
            center.pch = F,
            col = 'grey70',
            fill = T,
            fill.alpha = 0.5,
            add = T, plot.points = F,
            levels = 0.5,
            lwd = 1)

## June ellipse
dataEllipse(x = puntseq.HCO3_Na_vs_Ca_Na[11:20,2],
            y = puntseq.HCO3_Na_vs_Ca_Na[11:20,1],
            center.pch = F,
            col = 'grey70',
            fill = T,
            fill.alpha = 0.5,
            add = T, plot.points = F,
            levels = 0.5,
            lwd = 1)

## August ellipse and data points
dataEllipse(x = puntseq.HCO3_Na_vs_Ca_Na[21:30,2],
            y = puntseq.HCO3_Na_vs_Ca_Na[21:30,1],
            center.pch = F,
            col = 'grey70',
            fill = T,
            fill.alpha = 0.5,
            add = T, plot.points = F,
            levels = 0.5,
            lwd = 1)
points(y = puntseq.HCO3_Na_vs_Ca_Na[21:30,1], x = puntseq.HCO3_Na_vs_Ca_Na[21:30,2], pch = 21, 
       cex = 1.2, col = 'black',
       bg = c(alpha('#2157A4', transp), alpha('#3694D1', transp), alpha('#65C6E8', transp), 
              alpha('#9DD7ED', transp), alpha('#F7EC73', transp), alpha('#FDCB44', transp), 
              alpha('#F1861E', transp), alpha('#E63A11', transp), alpha('#D61015', transp), alpha('#D61015', transp)))

## legend
legend("bottomright", 
       cex = 1.3, pch = 21, 
       bty = 'n',
       col = 'black',
       title = 'August',
       title.col = 'black',
       legend = c('1', '2', '3', '4', '5', '6', '7', '8', '9'), 
       text.col = c('#2157A4', '#3694D1', '#65C6E8', '#9DD7ED', '#D6DE4F', 
                    '#FDCB44', '#F1861E', '#E63A11', '#a50026'), 
       pt.bg = c('#2157A4', '#3694D1', '#65C6E8', '#9DD7ED', '#D6DE4F', 
               '#FDCB44', '#F1861E', '#E63A11', '#a50026'))
dev.off()

## Month-by-month
pdf(str_c(outfolder,"Hydrochemistry/",date,"_Carbonation_world_rivers_individual_months.pdf"), width = 14, height = 8)
transp <- 0.7
par(mar = c(6,5,5,2))

# April
plot(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,1], x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,2], log = 'xy', 
     xlim = c(0.1, 100), ylim = c(0.1, 100), pch = 21, bg = 'black', col = 'white', ylab = "HCO3/Na", xlab = "Ca/Na",
     cex.lab = 1.5, main = "World River Carbonation Loads - vs. Cam April", cex.main = 3, cex = 2)
points(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,1], x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,2],
       cex = 2, pch = 21, bg = 'white', col = 'black')
points(y = puntseq.HCO3_Na_vs_Ca_Na[1:10,1], x = puntseq.HCO3_Na_vs_Ca_Na[1:10,2], pch = 24, cex = 2,
       bg = c(alpha('#2157A4', transp), alpha('#3694D1', transp), alpha('#65C6E8', transp), alpha('#9DD7ED', transp), alpha('#F7EC73', transp), 
              alpha('#FDCB44', transp), alpha('#F1861E', transp), alpha('#E63A11', transp), alpha('#D61015', transp), alpha('#D61015', transp)))
legend("topright", c('1 (Byrons Pool)', '2 (Byrons Pool)', '3 (Grantchester Meadows)', '4 (Granta)', '5 (Bridge of Sighs)', 
                     '6 (Fort St. George)', '7 (Green Dragon)', '8 (Sewage Pipe)', '9.1 (Baits Bite Lock)', '9.2  (Baits Bite Lock)'), 
       text.col = c('#2157A4','#3694D1','#65C6E8', '#9DD7ED', '#F7EC73', '#FDCB44', '#F1861E', '#E63A11', '#D61015', '#D61015'), bty = 'n')

# June
plot(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,1], x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,2], log = 'xy', 
     xlim = c(0.1, 100), ylim = c(0.1, 100), pch = 21, bg = 'black', col = 'white', ylab = "HCO3/Na", xlab = "Ca/Na",
     cex.lab = 1.5, main = "World River Carbonation Loads - vs. Cam June", cex.main = 3, cex = 2)
points(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,1], x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,2],
       cex = 2, pch = 21, bg = 'white', col = 'black')
points(y = puntseq.HCO3_Na_vs_Ca_Na[11:20,1], x = puntseq.HCO3_Na_vs_Ca_Na[11:20,2], pch = 23, cex = 2,
       bg = c(alpha('#2157A4', transp), alpha('#3694D1', transp), alpha('#65C6E8', transp), alpha('#9DD7ED', transp), alpha('#F7EC73', transp), 
              alpha('#FDCB44', transp), alpha('#F1861E', transp), alpha('#E63A11', transp), alpha('#D61015', transp), alpha('#D61015', transp)))
legend("topright", c('1 (Byrons Pool)', '2 (Byrons Pool)', '3 (Grantchester Meadows)', '4 (Granta)', '5 (Bridge of Sighs)', 
                     '6 (Fort St. George)', '7 (Green Dragon)', '8 (Sewage Pipe)', '9.1 (Baits Bite Lock)', '9.2  (Baits Bite Lock)'), 
       text.col = c('#2157A4','#3694D1','#65C6E8', '#9DD7ED', '#F7EC73', '#FDCB44', '#F1861E', '#E63A11', '#D61015', '#D61015'), bty = 'n')

# August
plot(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,1], x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,2], log = 'xy', 
     xlim = c(0.1, 100), ylim = c(0.1, 100), pch = 21, bg = 'black', col = 'white', ylab = "HCO3/Na", xlab = "Ca/Na",
     cex.lab = 1.5, main = "World River Carbonation Loads - vs. Cam August", cex.main = 3, cex = 2)
points(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,1], x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,2],
       cex = 2, pch = 21, bg = 'white', col = 'black')
points(y = puntseq.HCO3_Na_vs_Ca_Na[21:30,1], x = puntseq.HCO3_Na_vs_Ca_Na[21:30,2], pch = 25, cex = 2,
       bg = c(alpha('#2157A4', transp), alpha('#3694D1', transp), alpha('#65C6E8', transp), alpha('#9DD7ED', transp), alpha('#F7EC73', transp),
              alpha('#FDCB44', transp), alpha('#F1861E', transp), alpha('#E63A11', transp), alpha('#D61015', transp), alpha('#D61015', transp)))
legend("topright", c('1 (Byrons Pool)', '2 (Byrons Pool)', '3 (Grantchester Meadows)', '4 (Granta)', '5 (Bridge of Sighs)', 
                     '6 (Fort St. George)', '7 (Green Dragon)', '8 (Sewage Pipe)', '9.1 (Baits Bite Lock)', '9.2  (Baits Bite Lock)'), 
       text.col = c('#2157A4','#3694D1','#65C6E8', '#9DD7ED', '#F7EC73', '#FDCB44', '#F1861E', '#E63A11', '#D61015', '#D61015'), bty = 'n')

# Stacked
plot(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,1], x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] < 500,2], log = 'xy', 
     xlim = c(0.1, 100), ylim = c(0.1, 100), pch = 21, bg = 'black', col = 'white', ylab = "HCO3/Na", xlab = "Ca/Na",
     cex.lab = 1.5, main = "World River Carbonation Loads - vs. Cam", cex.main = 3, cex = 2)
points(y = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,1], x = gaillardet.HCO3_Na_vs_Ca_Na[gaillardet.HCO3_Na_vs_Ca_Na[,3] >= 500,2],
       cex = 2, pch = 21, bg = 'white', col = 'black')
points(y = puntseq.HCO3_Na_vs_Ca_Na[1:10,1], x = puntseq.HCO3_Na_vs_Ca_Na[1:10,2], pch = 24, cex = 1,
       bg = c(alpha('#2157A4', transp), alpha('#3694D1', transp), alpha('#65C6E8', transp), alpha('#9DD7ED', transp), alpha('#F7EC73', transp), 
              alpha('#FDCB44', transp), alpha('#F1861E', transp), alpha('#E63A11', transp), alpha('#D61015', transp), alpha('#D61015', transp)))
points(y = puntseq.HCO3_Na_vs_Ca_Na[11:20,1], x = puntseq.HCO3_Na_vs_Ca_Na[11:20,2], pch = 23, cex = 1,
       bg = c(alpha('#2157A4', transp), alpha('#3694D1', transp), alpha('#65C6E8', transp), alpha('#9DD7ED', transp), alpha('#F7EC73', transp), 
              alpha('#FDCB44', transp), alpha('#F1861E', transp), alpha('#E63A11', transp), alpha('#D61015', transp), alpha('#D61015', transp)))
points(y = puntseq.HCO3_Na_vs_Ca_Na[21:30,1], x = puntseq.HCO3_Na_vs_Ca_Na[21:30,2], pch = 25, cex = 1,
       bg = c(alpha('#2157A4', transp), alpha('#3694D1', transp), alpha('#65C6E8', transp), alpha('#9DD7ED', transp), alpha('#F7EC73', transp),
              alpha('#FDCB44', transp), alpha('#F1861E', transp), alpha('#E63A11', transp), alpha('#D61015', transp), alpha('#D61015', transp)))
legend("topright", c('1 (Byrons Pool)', '2 (Byrons Pool)', '3 (Grantchester Meadows)', '4 (Granta)', '5 (Bridge of Sighs)', 
                     '6 (Fort St. George)', '7 (Green Dragon)', '8 (Sewage Pipe)', '9.1 (Baits Bite Lock)', '9.2  (Baits Bite Lock)'), 
       text.col = c('#2157A4','#3694D1','#65C6E8', '#9DD7ED', '#F7EC73', '#FDCB44', '#F1861E', '#E63A11', '#D61015', '#D61015'), bty = 'n')
dev.off()
```

### Classifier comparison
#### On mock community 
##### Observed vs expected for Minimap2
```{r}
# create output directory
system(str_c("mkdir ",outfolder,"Tool_benchmarking"))

# Load marority vote classifiered set of 10,000 reads from sample 8
data <- read_tsv(str_c(outfolder,"Supplementary_data/2020-01-31_Minimap2_results_family_level_observed_vs_expected_full.tsv"))

ggscatter(data, x = "abundance_expected", y = "abundance_observed", size = 4)+
  geom_abline(intercept = 0, slope = 1, color="black", 
                 linetype="solid", size=1)+
  #geom_point(aes(color=Time))+
  scale_color_grey()+
  labs(title = "", caption = waiver(), x="expected abundance [%]", y="observed abundance [%]")+
  theme(legend.position="bottom")+
  xlim(0, 65)+
  ylim(0, 65)

ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_observed_vs-expected-plot_full.png"), width = 8, height = 8, units = "cm")
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_observed_vs-expected-plot_full.pdf"), width = 8, height = 8, units = "cm")

data.raw <- read_tsv(str_c(outfolder,"Supplementary_data/2020-01-31_Minimap2_results_family_level_observed_vs_expected_withoutEntero.tsv"))

ggscatter(data.raw, x = "abundance_expected", y = "abundance_observed", size = 4)+
  geom_abline(intercept = 0, slope = 1, color="black", 
                 linetype="solid", size=1)+
  #geom_point(aes(color=Time))+
  scale_color_grey()+
  labs(title = "", caption = waiver(), x="expected abundance [%]", y="observed abundance [%]")+
  theme(legend.position="bottom")+
  xlim(0, 35)+
  ylim(0, 35)

ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_observed_vs-expected-plot.png"), width = 8, height = 8, units = "cm")
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_observed_vs-expected-plot.pdf"), width = 8, height = 8, units = "cm") 
```

##### RMSE comparision
```{r}
# 1. Load table
###############

## Cambridge weather sensor data
tool.performance <- as.matrix(read_xlsx(str_c(outfolder,"Supplementary_data/Fig_S1_b_table.xlsx"), sheet = 1))

# 2. Plot
#########

# version 1
results <- as.numeric(tool.performance[nrow(tool.performance),-1])
names(results) <- colnames(tool.performance)[-1]

pdf(str_c(outfolder,"Tool_benchmarking/",date,"_RMSE-.pdf"), width = 7, height = 6)
plot(x = results[-1], y = rep(1, length(results[-1])),
     pch = 21, cex = 2, bg = 'black', col = 'white', bty = 'n',
     xlab = '', ylab = '', main = '', xaxt = 'n', yaxt = 'n',
     ylim = c(0.75,2), xlim = c(0, 35))
axis(1, at = seq(f = 0, t = 35, by = 5), labels = seq(f = 0, t = 35, by = 5), 
     cex.axis = 2, padj = 0.5)
lines(x = c(0, 35), y = c(1,1))
for (i in 2:length(results)){
  lines(x = c(results[i], results[i]), y = c(1, 1.5), lwd = 0.5, lty = 1)
}
dev.off()

# version 2
df <- data.frame(RMSE_reduced=c(0.00,6.52,6.57,6.75,6.78,6.88,6.95,7.04,7.53,7.65,9.38,11.42,11.78,16.58),
Classifier=c("expected","IDTAXA","Minimap2 (k = 15)","Blastn","SPINGO (k = 8)","SINTAX","Megablast","MAPseq","QUIIME 2 Blastn","QUIIME 2 Sklearn","Mothur (k = 6)","Kraken2 (k = 21)",	"Centrifuge",	"RDP"))

set.seed(42)

ggplot(df, aes(x = RMSE_reduced, y = 1, label = Classifier)) +
  geom_point(aes(colour="white", fill="black"), shape = 21,size = 4, colour = "white")+
  geom_text_repel(
    nudge_y      = 0.03,
    direction    = "x",
    angle        = 90,
    vjust        = 0,
    segment.size = 0.5
  ) +
  xlim(0, 16.58) +
  ylim(1, 0.8) +
  theme_classic() +
  theme(
    axis.line.y  = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.title.y = element_blank()
  )+
  theme(legend.position="none")

ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_line-plot.png"), width = 15, height = 8, units = "cm")
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_line-plot.pdf"), width = 15, height = 8, units = "cm") 
```

#### On freshwater community (sample 8)
##### Comparison versus mayority vote
```{r}
# loop over all three sets
for (set in 1:3){

  # Load marority vote classifiered set of 10,000 reads from sample 8
  mayority.vote.raw <- read_tsv(str_c(outfolder,"Supplementary_data/summary_sample8_rep",set,"_family_majority_vote.tsv"))
  
  # count number of classified and unclassified reads on family level
  for (i in 2:(ncol(mayority.vote.raw))){
    classification.col <- as.data.frame(table(!is.na(mayority.vote.raw[,i])))
    classification.col <- mutate(classification.col, Tool = colnames(mayority.vote.raw[,i])) %>%
      mutate(Perc = (Freq/nrow(mayority.vote.raw)*100)) %>%
      rename(Type = Var1) %>%
      mutate(Type = str_replace(Type, pattern = "FALSE", "unclassified")) %>%
      mutate(Type = str_replace(Type, pattern = "TRUE", "classified")) %>%
      mutate(Set = set) 
        
    if (i == 2){
      classification.total <- classification.col
    } else {
      classification.total <- rbind(classification.total, classification.col)
    }
  }
  
  mayority.vote <- mayority.vote.raw
  # loop over all rows and columns and compare against mayority vote (in last column) -> create binary matrix
  for (i in 2:(ncol(mayority.vote)-1)){
    for (j in 1:nrow(mayority.vote.raw)){
      if (!is.na(mayority.vote[j,ncol(mayority.vote)])){
        if (!is.na(mayority.vote[j,i])){
          if ( mayority.vote[j,i] == mayority.vote[j,ncol(mayority.vote)]){
            mayority.vote[j,i] = "3" # Match
          } 
          else {
            mayority.vote[j,i] = "2" # Mismatch
          }
        } 
        else {
            mayority.vote[j,i] = "2" # Mismatch
        }
      } 
      else {
        if (!is.na(mayority.vote[j,i])){
          mayority.vote[j,i] = "1" # Over classification
        } else {
          mayority.vote[j,i] = "0" 
        }
      }
    }
  }
  
  # add to classification.total 
  for (i in 2:(ncol(mayority.vote)-1)){
    classification.all <- as.data.frame(table(mayority.vote[,i]))
    classification.all <- mutate(classification.all, Tool = colnames(mayority.vote[,i])) %>%
      mutate(Perc = (Freq/nrow(mayority.vote.raw)*100)) %>%
      rename(Type = Var1) %>%
      mutate(Type = str_replace(Type, pattern = "1", "over classified")) %>%
      mutate(Type = str_replace(Type, pattern = "2", "false classified")) %>%
      mutate(Type = str_replace(Type, pattern = "3", "true classified")) %>%
      mutate(Set = set) %>%
      filter(Type != "0")
    
    classification.total <- rbind(classification.total, classification.all)
  }

  # add missing value
  Type <- c("over classified")
  Tool <- c("minimap2(12)")
  Perc <- c(0)
  Freq <- c(0)
  Set <- set
  classification.minimap <- data.frame(Type, Freq, Tool, Perc,Set)
  classification.total.final <- rbind(classification.total, classification.minimap)
  
  classification.total.final <- mutate(classification.total.final, Panel = ifelse(Type == "classified" | Type == "unclassified", "overview", "classified reads")) %>%
    transform(Panel = factor(Panel, levels = c("overview","classified reads")))%>%
    transform(Type = factor(Type, levels = c("classified","true classified","false classified","over classified","unclassified")))
    
  # plot data
  ggplot(filter(classification.total.final,Type != "classified" & Tool != "gt" & Tool != "qiime_sklearn"), aes(x = fct_reorder(Tool,Perc), y = Perc, fill = Type)) +
    geom_bar(stat="identity", position = position_dodge(), width = 0.7,  color="grey30")+
      geom_hline(yintercept=100, linetype="dashed", color = "grey")+
        labs( y = "Percentage", x="")+
        #facet_grid(rows = vars(Panel))+ 
        scale_y_continuous(breaks=seq(0, 108, 20), limits = c(0,108))+
        theme_bw()+
        theme(panel.grid.major.x = element_blank())+
        theme(plot.title = element_text(hjust = 0.5, size = 11))+
        theme(axis.text.x = element_text(size = 8))+
        theme(axis.text.y = element_text(size = 7))+
        theme(legend.text = element_text(size = 8))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
        theme(legend.title = element_blank())+
        scale_fill_npg()+
        #scale_fill_manual(values = c("#56B4E9","blue","red","grey40","grey85"))+
      geom_text(aes(label = round(Perc,0), x = fct_reorder(Tool,Perc), y = Perc), position = position_dodge(width = 0.3), vjust = -0.6, color="black", size=3)+
    theme(legend.position="top")
  
  ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_sample8_mv-comparision_rep",set,".png"), width = 12, height = 8, units = "cm")
  ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_sample8_mv-comparision_rep",set,".pdf"), width = 12, height = 8, units = "cm") 

}  
```

##### Relative abundance of freshwater sample for all classifier
```{r}
# loop over all three sets
for (set in 1:3){

  # Load marority vote classifiered set of 10,000 reads from sample 8
  mayority.vote.raw <- read_tsv(str_c(outfolder,"Supplementary_data/summary_sample8_rep",set,"_family_majority_vote.tsv"))
  
  for (i in 2:(ncol(mayority.vote.raw))){
    abundance.col <- as.data.frame(table(mayority.vote.raw[,i]))
    abundance.col <- mutate(abundance.col, Tool = colnames(mayority.vote.raw[,i])) %>%
      mutate(Perc = (Freq/nrow(mayority.vote.raw)*100)) %>%
      rename(Family = Var1) #%>%
      #mutate(Type = str_replace(Type, pattern = "FALSE", "unclassified")) %>%
      #mutate(Type = str_replace(Type, pattern = "TRUE", "classified"))
    
    if (i == 2){
      abundance.total <- abundance.col
    } else {
       abundance.total <- rbind(abundance.total, abundance.col)
    }
  }
  
  # select top 10 most abundant taxa with the highest total abundance (percentage) among all samples
  subset_tmp <- abundance.total %>%
    group_by(Family) %>%
    summarize(Sum = sum(Freq)) %>%
    ungroup() %>%
    arrange(desc(Sum)) 
  
  legend <- subset_tmp$Family %>%
    unique()
  a<- as.character(legend[1:10])
  legend <- c(a, "Other")
  
  subset_rank_top10 <- transform(abundance.total, Family = factor(Family, levels = legend)) %>%
          mutate(Tool = str_replace(Tool, pattern = "gt", "majority vote")) %>%
          transform(abundance.total, Tool = factor(Tool, levels = c("majority vote","spingo(8)","minimap2(12)","idtaxa","mapseq","megablast","blastn","qiime_blastn","sintax","kraken2","centrifuge","rdp","mothur(6)","qiime_sklearn"))) %>%
          mutate(Family = replace_na(Family, "Other"))
  
  
  # plot relativ abundance (Version 3) the top 10 most abundant taxa with the highest total abundance (percentage) among all samples
  ggplot(data = filter(subset_rank_top10, Tool != "qiime_sklearn"), mapping = aes(x = Tool, y = Perc, fill = forcats::fct_rev(Family)))+
    geom_bar(stat="identity", position = "fill", width = 0.3)+
    #facet_grid(cols = vars(Location))+ 
    labs(y = "Abundance")+
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5, size = 22))+
    theme(axis.text.x = element_text(size = 16))+
    theme(axis.text.y = element_text(size = 14))+
    theme(legend.text = element_text(size = 16))+
    theme(legend.title = element_blank())+
    scale_fill_manual(values = rev(brewer.pal(11,"Set3")))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    theme(legend.position="none")
  
  # save plot
  ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_Rel_abundance_sample8_run",set,".png"), width = 13, height = 10, units = "cm")
  ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_Rel_abundance_sample8_run",set,".pdf"), width = 13, height = 10, units = "cm")
}

```

#### Runtime and Memory usage
```{r}

## speed
speed <- read_xlsx(str_c(outfolder,"Supplementary_data/2020-11-03_tool_benchmark_resources.xlsx"), sheet = 1) %>%
  mutate(Set = str_replace(Set, pattern = "set", "")) %>%
  separate(Set, c("Reads","Rep"), "_") %>%
  gather(key = "Tool", value = "Speed", 3:14)

speed_av_1 <- speed %>%
  group_by(Tool,Reads) %>%
  summarize(Speed_av = mean(Speed, na.rm = TRUE)) %>%
  ungroup()
speed_av_2 <- speed %>%
  group_by(Tool,Reads) %>%
  summarize(Speed_sd = sd(Speed, na.rm = TRUE)) %>%
  ungroup()
speed_av <- left_join(speed_av_1, speed_av_2) 

order.df <- speed_av %>%
  group_by(Tool) %>%
  summarize(Mean = mean(Speed_av, na.rm = TRUE)) %>%
  arrange(Mean) %>%
  select(Tool) %>%
  unique()
order <- order.df$Tool

speed$Tool <- factor(speed$Tool, levels = order)
speed$Reads <- factor(speed$Reads, levels = c("100","1000","10000")) 
  
## average memory
avmem <- read_xlsx(str_c(outfolder,"Supplementary_data/2020-11-03_tool_benchmark_resources.xlsx"), sheet = 2) %>%
  mutate(Set = str_replace(Set, pattern = "set", "")) %>%
  separate(Set, c("Reads","Rep"), "_") %>%
  gather(key = "Tool", value = "Mem", 3:14) %>%
  mutate(Mem = Mem/1000)
avmem$Tool <- factor(avmem$Tool, levels = order) 
avmem$Reads <- factor(avmem$Reads, levels = c("100","1000","10000")) 

# version 1
# plot runtime
ggline(speed, x = "Reads", y = "Speed", add = "mean_sd", color = "Tool", size = 0.5, shape="Tool") +
  scale_y_continuous(trans='log10', breaks = trans_breaks("log10", function(x) 10^x),labels = trans_format("log10", math_format(10^.x)))+scale_fill_grey() + 
  scale_color_grey() + 
  theme_classic() +
  xlab("Reads") +
  ylab("Runtime (s)") +
  guides(colour = guide_legend(reverse=T)) +
  guides(shape = guide_legend(reverse=T))

# save plot
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_runtime.pdf"), width = 11, height = 10, units = "cm")
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_runtime.png"), width = 11, height = 10, units = "cm")

# plot mem
ggline(avmem, x = "Reads", y = "Mem", add = "mean_sd", color = "Tool", size = 0.5, shape="Tool") +
  scale_color_grey() + 
  theme_classic() +
  xlab("Reads") +
  ylab("Average memory use (Gb)") +
  guides(colour = guide_legend(reverse=T)) +
  guides(shape = guide_legend(reverse=T))

# save plot
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_memory.pdf"), width = 11, height = 10, units = "cm")
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_memory.png"), width = 11, height = 10, units = "cm")

# version 2
# plot runtime
ggplot(speed, aes(x=Tool, y=Speed, color=Reads))+
  geom_beeswarm(data=filter(speed, Reads =="100"), alpha = 0.8, priority='ascending')+
  geom_beeswarm(data=filter(speed, Reads =="1000"), alpha = 0.8, priority='ascending')+
  geom_beeswarm(data=filter(speed, Reads =="10000"), alpha = 0.8, priority='ascending')+
  scale_y_continuous(trans='log10', breaks = trans_breaks("log10", function(x) 10^x),labels = trans_format("log10", math_format(10^.x)))+
  scale_color_grey() +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust = 1))+
  xlab("") +
  ylab("Runtime (s)")

# save plot
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_runtime_v2.pdf"), width = 20, height = 8, units = "cm")
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_runtime_v2.png"), width = 20, height = 8, units = "cm")

# plot mem
ggplot(avmem, aes(x=Tool, y=Mem, color=Reads))+
  geom_beeswarm(data=filter(avmem, Reads =="100"), alpha = 0.8, priority='ascending')+
  geom_beeswarm(data=filter(avmem, Reads =="1000"), alpha = 0.8, priority='ascending')+
  geom_beeswarm(data=filter(avmem, Reads =="10000"), alpha = 0.8, priority='ascending')+
  scale_color_grey() +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust = 1))+
  xlab("") +
  ylab("Average memory use (Gb)") 

# save plot
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_memory_v2.pdf"), width = 20, height = 8, units = "cm")
ggsave(str_c(outfolder,"Tool_benchmarking/",date,"_memory_v2.png"), width = 20, height = 8, units = "cm")

```
